### 2021-12-11 paper_replication.Rmd
Basically just a brief TODO list so far.

---
title: "paper_replication"
output: pdf_document
---
TODO: classify counties as highly restrictive or not
    - option 1: we can use the summary data that Alice linked to
    (https://www.guttmacher.org/article/2019/08/state-abortion-policy-landscape-hostile-supportive)
    - option 2: dig through the detailed legislation data and use the 3/4 types classification they did in the paper (http://lawatlas.org/datasets/abortion-laws)
    I'd say let's start with the summary data, and we do the extra work to make a more faithful replication later if we have time at the end.

TODO: Calculate distances to nearest clinic
    - I added 'county_pop_centroids.csv' from the Census Bureau
    - In the paper they did centroid of county of residence centroid of nearest county  with a clinic, but I think because we have the lat and long in the provider list we can do centroid of county of residence to nearest clinic

TODO: model the propensity weights
    - If I understand this right, we're modeling the propensity for a county to be located in a highly restrictive state.  Is that right?  So we want an estimated probability of a binary outcome, so we want logistic regression. Is that right?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
options(tidyverse.quiet=TRUE)
library(tidyverse)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
provider_list_geo <- read.csv('data/provider_list_geo.csv')

```

\newpage

## Sources

Data on the county population centroids came from the United States Census Bureau.
Available at https://www.census.gov/geographies/reference-files/time-series/geo/centers-population.html Accessed 12/11/21.

## Code Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

### 2021-12-11 DataCleaning.R
Add files via upload

@@ -1,0 +1,116 @@
+library(tidyverse)
+library(readxl)
+library(dplyr)
+
+providers <- read.csv("provider_list_geo.csv")
+countycentroids <- read.csv("county_pop_centroids.csv")
+countydem <- read.csv("county_demographics.csv")
+abortions <- read.csv("abortions_by_county.csv")
+
+preabortionreqs <- read_excel("PreabortionRequirements.xlsx", sheet = 2)
+preabortionreqs <- preabortionreqs %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "MWP_WaitPeriod")
+preabortionreqs$MWP_WaitPeriod[preabortionreqs$MWP_WaitPeriod == "."] <- 0
+preabortionreqs <- preabortionreqs %>% dplyr::rename(waitperiod = MWP_WaitPeriod)
+
+minorsreqs <- read_excel("Minors.xlsx", sheet = 2)
+minorsreqs <- minorsreqs %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "Minor_consentreq")
+minorsreqs$Minor_consentreq[minorsreqs$Minor_consentreq == "."] <- 0
+minorsreqs <- minorsreqs %>% dplyr::rename(parentalconsent = Minor_consentreq)
+
+abortionbans <- read_excel("AbortionBans.xlsx", sheet = 2)
+abortionbans <- abortionbans %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "Prohibit_Req")
+abortionbans$Prohibit_Req[abortionbans$Prohibit_Req == "."] <- 0
+abortionbans <- abortionbans %>% dplyr::rename(gestcutoff = Prohibit_Req)
+
+traphospital <- read_excel("TRAPHospitalization.xlsx", sheet = 2)
+traphospital <- traphospital %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "trap-hr-reghosp")
+traphospital <- traphospital %>% dplyr::rename(traphospitalreqs = "trap-hr-reghosp")
+
+trapasc <- read_excel("TRAPASC.xlsx", sheet = 2)
+trapasc <- trapasc %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "trap-ascr-regambsurgcenrequirm")
[…]

### 2021-12-11 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
+
+BIG PROBLEM: data on abortion restrictions starts at the end of December 2018, but data on abortion counts only goes up to (max) 2018
+   * Option 1: Somehow find abortion restriction data from before December 2018 - I tried with no luck yet
+   * Option 2: Somehow find abortion count data for 2019 on - Also tried with no luck yet 
+         - I found this by state but not by county.
+
 TODO: classify counties as highly restrictive or not
@@
     - In the paper they did centroid of county of residence centroid of nearest county  with a clinic, but I think because we have the lat and long in the provider list we can do centroid of county of residence to nearest clinic
+
+TODO: Calculate abortion rates by county
 

### 2021-12-12 DataCleaning.R
Update DataCleaning.R

@@ -6,6 +6,10 @@ providers <- read.csv("provider_list_geo.csv")
 countycentroids <- read.csv("county_pop_centroids.csv")
 countydem <- read.csv("county_demographics.csv")
 abortions <- read.csv("abortions_by_county.csv")
+names(abortions)
+aggregate(X2017 ~ state, data=abortions, function(x) {sum(is.na(x))/length(x)}, na.action = NULL)
+aggregate(X2018 ~ state, data=abortiondata, function(x) {sum(is.na(x))/length(x)}, na.action = NULL)
+
 
 preabortionreqs <- read_excel("PreabortionRequirements.xlsx", sheet = 2)
 preabortionreqs <- preabortionreqs %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "MWP_WaitPeriod")
@@ -43,71 +47,7 @@ min(traphospital$"Effective Date")
 min(trapasc$"Effective Date")
 min(trapafl$"Effective Date")
 
-#####state data
-arizona <- read.csv("arizona.csv")
-arizona$state <- "Arizona"
-delaware <- read.csv("delaware.csv")
-delaware$state <- "Delaware"
-abortiondata <- full_join(arizona, delaware, by = names(delaware))
-georgia <- read.csv("georgia.csv")
-georgia$state <- "Georgia"
-abortiondata <- full_join(abortiondata, georgia, by = names(georgia))
-illinois <- read.csv("illinois.csv")
-illinois$state <- "Illinois"
-abortiondata <- full_join(abortiondata, illinois, by = names(illinois))
-indiana <- read.csv("indiana.csv")
-indiana$state <- "Indiana"
-abortiondata <- full_join(abortiondata, indiana, by = names(indiana))
-michigan <- read.csv("michigan.csv")
-michigan$state <- "Michigan"
-abortiondata <- full_join(abortiondata, michigan, by = names(michigan))
-newyork <- read.csv("new_york.csv")
-newyork$state <- "New York"
-abortiondata <- full_join(abortiondata, newyork, by = names(newyork))
-northcarolina <- read.csv("north_carolina.csv")
-northcarolina$state <- "North Carolina"
-abortiondata <- full_join(abortiondata, northcarolina, by = names(northcarolina))
-ohio <- read.csv("ohio.csv")
-ohio$state <- "Ohio"
-abortiondata <- full_join(abortiondata, ohio, by = names(ohio))
[…]

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -54,3 +54,13 @@ write.csv(abortionbans, "abortionbans.csv")
 write.csv(traphospital, "traphospital.csv")
 write.csv(trapasc, "trapasc.csv")
 write.csv(trapafl, "trapafl.csv")
+
+
+state_restrictions <- data.frame(state = state.name)
+
+state_restrictions$parental_involvement <- c(1, 0, 1, 1, 0, 1, 0, rep(1, 3), 0, rep(1, 7), 0, rep(1, 6), 0, 1, rep(0, 5), rep(1, 4), 0, rep(1, 7), 0, 1, 0, rep(1, 3))
+state_restrictions$two_trips <- c(0, 0, 1, rep(0, 10), 1, rep(0, 3), 1, rep(0, 5), 1, 1, rep(0, 9), 1, rep(0, 8), 1, rep(0, 4), 1, 0)
+state_restrictions$trap <- c(1, 0, 0, 1, rep(0, 9), 1, 0, 0, 1, 1, 0, 0, 0, 1, 0, 1, 1, rep(0,9), 1, 1, 0, 1, 0, 1, 1, 0, 0, 1, rep(0, 6))
+state_restrictions$unconstitutional_ban <- c(1, rep(0, 7), 1, 1, 0, 1, 0, 1, 1, rep(0, 5), 1, 1, rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 4), 0, 1, 1, 0, 1, 0, 0, 0,1) 
+
+write.csv(state_restrictions, "state_restrictions2010")

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -64,3 +64,15 @@ state_restrictions$trap <- c(1, 0, 0, 1, rep(0, 9), 1, 0, 0, 1, 1, 0, 0, 0, 1, 0
 state_restrictions$unconstitutional_ban <- c(1, rep(0, 7), 1, 1, 0, 1, 0, 1, 1, rep(0, 5), 1, 1, rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 4), 0, 1, 1, 0, 1, 0, 0, 0,1) 
 
 write.csv(state_restrictions, "state_restrictions2010")
+
+library(geosphere)
+
+for (i in 1:nrow(countycentroids)) {
+  lat <- countycentroids$LATITUDE[i]
+  long <- countycentroids$LONGITUDE[i]
+  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))
+  countycentroids$distancetoprovider[i] <- min(dists)
+}
+
+write.csv(countycentroids, "countypopcentroids")
+

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -76,3 +76,40 @@ for (i in 1:nrow(countycentroids)) {
 
 write.csv(countycentroids, "countypopcentroids")
 
+
+abortions <- read.csv("cleanedabortiondata.csv")
+names(abortions)
+abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
+
+countydem <- read.csv("county_demographics.csv")
+library(stringr)
+for (i in 1:nrow(countydem)) {
+  countydem$state[i] <- unlist(strsplit(countydem$NAME[i], ", "))[length(unlist(strsplit(countydem$NAME[i], ", ")))]
+  county <- unlist(strsplit(countydem$NAME[i], " "))[unlist(strsplit(countydem$NAME[i], " ")) != "County," & unlist(strsplit(countydem$NAME[i], " ")) != countydem$state[i]]
+  countydem$county[i] <- paste(county, collapse = " ")
+}
+
+countydem <- countydem %>% filter(countydem$state %in% state.name)
+
+for (i in 1:nrow(countydem)) {
+  countydem$state.abb[i] <- state.abb[which(state.name == countydem$state[i])]
+}
+
+for (i in 1:nrow(abortions2010)) {
+  if (length(countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]) == 0) {
+    abortions2010$Women[i] <- NA
+  }
+  else {
+    abortions2010$Women[i] <- countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]
+  }
+}
+
+abortions2010$X2010 <- na_if(abortions2010$X2010, "?")
+abortions2010$X2010 <- na_if(abortions2010$X2010, "")
+abortions2010$X2010 <- as.numeric(abortions2010$X2010)
+abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)
+
+
+write.csv(abortions2010, "abortionrates2010")
+
+

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -1,16 +1,19 @@
+#load required libraries
 library(tidyverse)
 library(readxl)
 library(dplyr)
 
+#load in data
 providers <- read.csv("provider_list_geo.csv")
 countycentroids <- read.csv("county_pop_centroids.csv")
 countydem <- read.csv("county_demographics.csv")
 abortions <- read.csv("abortions_by_county.csv")
-names(abortions)
+
+#look at missing data by state for a given year
 aggregate(X2017 ~ state, data=abortions, function(x) {sum(is.na(x))/length(x)}, na.action = NULL)
 aggregate(X2018 ~ state, data=abortiondata, function(x) {sum(is.na(x))/length(x)}, na.action = NULL)
 
-
+#load in specific abortion ban data
 preabortionreqs <- read_excel("PreabortionRequirements.xlsx", sheet = 2)
 preabortionreqs <- preabortionreqs %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "MWP_WaitPeriod")
 preabortionreqs$MWP_WaitPeriod[preabortionreqs$MWP_WaitPeriod == "."] <- 0
@@ -34,12 +37,12 @@ trapasc <- read_excel("TRAPASC.xlsx", sheet = 2)
 trapasc <- trapasc %>% dplyr::select("Jurisdiction", "Effective Date", "Valid Through Date", "trap-ascr-regambsurgcenrequirm")
 trapasc <- trapasc %>% dplyr::rename(trapascreqs = "trap-ascr-regambsurgcenrequirm")
 
-
 trapafl <- read_excel("TRAPAFL.xlsx", sheet = 2)
 trapafl <- trapafl %>% dplyr::select("Jurisdictions", "Effective Date", "Valid Through Date", "trap-aflr-regaborprocforaborprov2") %>% dplyr::rename(licensingreqs = "trap-aflr-regaborprocforaborprov2")
 
 trapafl$licensingreqs <- case_when(trapafl$licensingreqs == "No, but the law imposes requirements for transfer agreements and/or admitting privileges" | trapafl$licensingreqs == "Yes" ~ 1, trapafl$licensingreqs == "No" ~ 0)
 
+#look at years represented by specific ban data
 min(preabortionreqs$"Effective Date")
 min(minorsreqs$"Effective Date")
[…]

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -88,18 +88,49 @@ abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
 
 # clean countydem data to make counties and states match abortion count data 
 countydem <- read.csv("county_demographics.csv")
+
 library(stringr)
 for (i in 1:nrow(countydem)) {
   countydem$state[i] <- unlist(strsplit(countydem$NAME[i], ", "))[length(unlist(strsplit(countydem$NAME[i], ", ")))]
-  county <- unlist(strsplit(countydem$NAME[i], " "))[unlist(strsplit(countydem$NAME[i], " ")) != "County," & unlist(strsplit(countydem$NAME[i], " ")) != countydem$state[i]]
-  countydem$county[i] <- paste(county, collapse = " ")
 }
+
+countydem$county <- gsub(countydem$NAME, pattern = ", [[:print:]]*$", replacement = "")
+countydem$county <- gsub(countydem$county, pattern = " County", replacement = "")
 
 countydem <- countydem %>% filter(countydem$state %in% state.name)
 
 for (i in 1:nrow(countydem)) {
   countydem$state.abb[i] <- state.abb[which(state.name == countydem$state[i])]
 }
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "ind. city", replacement = "city")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Dekalb", replacement = "DeKalb")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Desoto", replacement = "DeSoto")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "City", replacement = "city")
+
+countydem$county <- gsub(countydem$county, pattern = "City", replacement = "city")
+
[…]

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -154,3 +154,21 @@ abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]
 abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)
 
 write.csv(abortions2010, "abortionrates2010")
+
+countypres <- read.csv("countypres.csv")
+countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
+countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)
+countypresdem2008$county_name[!(countypresdem2008$county_name %in% countydem$county)]
+
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "City", replacement = "city")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Dekalb", replacement = "DeKalb")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Desoto", replacement = "DeSoto")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
+
+countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
+names(countypresdem2008)
+countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)
+
+countydem <- left_join(countydem, countypresdem2008, by = c("state.abb", "county"))
+write.csv(countydem, "countydem.csv")
+

### 2021-12-13 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
-
-BIG PROBLEM: data on abortion restrictions starts at the end of December 2018, but data on abortion counts only goes up to (max) 2018
-   * Option 1: Somehow find abortion restriction data from before December 2018 - I tried with no luck yet
-   * Option 2: Somehow find abortion count data for 2019 on - Also tried with no luck yet 
-         - I found this by state but not by county.
 
@@
     I'd say let's start with the summary data, and we do the extra work to make a more faithful replication later if we have time at the end.
-
-TODO: Calculate distances to nearest clinic
-    - I added 'county_pop_centroids.csv' from the Census Bureau
-    - In the paper they did centroid of county of residence centroid of nearest county  with a clinic, but I think because we have the lat and long in the provider list we can do centroid of county of residence to nearest clinic
-
-TODO: Calculate abortion rates by county
 
@@
 provider_list_geo <- read.csv('data/provider_list_geo.csv')
-
 ```
@@
 
+Data on the county-level voting totals came from the Harvard Dataverse.
+Available at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ. Accessed 12/13/21.
+
 ## Code Appendix

### 2021-12-13 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#load packages
 options(tidyverse.quiet=TRUE)
 library(tidyverse)
+library(readxl)
+library(dplyr)
+library(geosphere)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-provider_list_geo <- read.csv('data/provider_list_geo.csv')
+#load in data
+providers <- read.csv("provider_list_geo.csv")
+countycentroids <- read.csv("county_pop_centroids.csv")
+countydem <- read.csv("county_demographics.csv")
+abortions <- read.csv("cleanedabortiondata.csv")
+restrictions <- read.csv("state_restrictions2010.csv")
+```
+
+```{r}
+########### calculate abortion rates in 2010 ###########
+abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
+
+# clean countydem data to make counties and states match abortion count data 
+countydem <- read.csv("county_demographics.csv")
+
+library(stringr)
+for (i in 1:nrow(countydem)) {
+  countydem$state[i] <- unlist(strsplit(countydem$NAME[i], ", "))[length(unlist(strsplit(countydem$NAME[i], ", ")))]
+}
+
+countydem$county <- gsub(countydem$NAME, pattern = ", [[:print:]]*$", replacement = "")
+countydem$county <- gsub(countydem$county, pattern = " County", replacement = "")
+
+countydem <- countydem %>% filter(countydem$state %in% state.name)
+
+for (i in 1:nrow(countydem)) {
+  countydem$state.abb[i] <- state.abb[which(state.name == countydem$state[i])]
+}
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "ind. city", replacement = "city")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Dekalb", replacement = "DeKalb")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Desoto", replacement = "DeSoto")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "City", replacement = "city")
+
+countydem$county <- gsub(countydem$county, pattern = "City", replacement = "city")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")
+
+abortions2010$county <- gsub(abortions2010$county, pattern = "Lamoure", replacement = "LaMoure")
+
+for (i in 1:nrow(abortions2010)) {
+  if (length(countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]) == 0) {
+    abortions2010$Women[i] <- NA
+  }
+  else {
+    abortions2010$Women[i] <- countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]
+  }
+}
+
+#fix NA values
+abortions2010$X2010 <- na_if(abortions2010$X2010, "?")
+abortions2010$X2010 <- na_if(abortions2010$X2010, "")
+abortions2010$X2010 <- as.numeric(abortions2010$X2010)
+
+abortions2010 <- abortions2010 %>% filter(!(county %in% c("NONRESIDENTS", "PROTECTED", "OTHER STATES", "OTHER COUNTRIES", "NON-RESIDENTS", "OTHER", "NEVADA PROTECTED", "NEVADA UNSPECIFIED", "NEBRASKA PROTECTED", "county", "MARYLAND", "St.
+Louis County")))
+
+abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]
+
+#make rates
+abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)
+
+countypres <- read.csv("countypres.csv")
+countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
+countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)
+countypresdem2008$county_name[!(countypresdem2008$county_name %in% countydem$county)]
+
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "City", replacement = "city")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Dekalb", replacement = "DeKalb")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Desoto", replacement = "DeSoto")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
+
+countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
+names(countypresdem2008)
+countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)
+
+countydem <- left_join(countydem, countypresdem2008, by = c("state.abb", "county"))
+```
+
+```{r}
+# calculate distances from county centroids to closest provider location
+
+for (i in 1:nrow(countycentroids)) {
+  lat <- countycentroids$LATITUDE[i]
+  long <- countycentroids$LONGITUDE[i]
+  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))
+  countycentroids$distancetoprovider[i] <- min(dists)
+}
+
 ```

### 2021-12-13 DataCleaning.R
Update DataCleaning.R

@@ -51,7 +51,6 @@ min(trapasc$"Effective Date")
 min(trapafl$"Effective Date")
 
 #update files
-write.csv(abortions, "cleanedabortiondata.csv")
 write.csv(preabortionreqs, "preabortionreqs.csv")
 write.csv(minorsreqs, "minorreqs.csv")
 write.csv(abortionbans, "abortionbans.csv")
@@ -67,7 +66,7 @@ state_restrictions$two_trips <- c(0, 0, 1, rep(0, 10), 1, rep(0, 3), 1, rep(0, 5
 state_restrictions$trap <- c(1, 0, 0, 1, rep(0, 9), 1, 0, 0, 1, 1, 0, 0, 0, 1, 0, 1, 1, rep(0,9), 1, 1, 0, 1, 0, 1, 1, 0, 0, 1, rep(0, 6))
 state_restrictions$unconstitutional_ban <- c(1, rep(0, 7), 1, 1, 0, 1, 0, 1, 1, rep(0, 5), 1, 1, rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 3), rep(0, 3), rep(1, 4), 0, 1, 1, 0, 1, 0, 0, 0,1) 
 
-write.csv(state_restrictions, "state_restrictions2010")
+write.csv(state_restrictions, "state_restrictions2010.csv")
 
 # calculate distances from county centroids to closest provider location
 library(geosphere)
@@ -81,6 +80,9 @@ for (i in 1:nrow(countycentroids)) {
 
 write.csv(countycentroids, "countypopcentroids")
 
+names(abortions)
+abortions <- abortions[, c(1:2, 46, 34:45, 22:31, 3:21, 32:33, 47)]
+write.csv(abortions, "cleanedabortiondata.csv")
 
 abortions <- read.csv("cleanedabortiondata.csv")
 names(abortions)

### 2021-12-13 paper_replication.Rmd
Update paper_replication.Rmd

@@
 library(geosphere)
+library(stringr)
 ```
@@
 
-library(stringr)
 for (i in 1:nrow(countydem)) {

### 2021-12-13 paper_replication.Rmd
Update paper_replication.Rmd

@@
 restrictions <- read.csv("state_restrictions2010.csv")
+countypres <- read.csv("countypres.csv")
 ```
@@
 
-countypres <- read.csv("countypres.csv")
+```
+
+```{r}
+## calculate percent democratic votes
 countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
@@
 
+```{r}
+#make new variable for high vs. low restrictions
+restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
+                                        highorlow = case_when(totalrestrict >= 3 ~ "high", 
+                                                              totalrestrict < 3 ~ "low"))
+```
+
+```{r}
+#make propensity scores
+
+```
+
 \newpage

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
-
-TODO: classify counties as highly restrictive or not
-    - option 1: we can use the summary data that Alice linked to
-    (https://www.guttmacher.org/article/2019/08/state-abortion-policy-landscape-hostile-supportive)
-    - option 2: dig through the detailed legislation data and use the 3/4 types classification they did in the paper (http://lawatlas.org/datasets/abortion-laws)
-    I'd say let's start with the summary data, and we do the extra work to make a more faithful replication later if we have time at the end.
 
@@
 abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "ind. city", replacement = "city")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Dekalb", replacement = "DeKalb")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Desoto", replacement = "DeSoto")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Hawai'i", replacement = "Hawaii")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "City", replacement = "city")
-
 countydem$county <- gsub(countydem$county, pattern = "City", replacement = "city")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")
-
 abortions2010$county <- gsub(abortions2010$county, pattern = "Lamoure", replacement = "LaMoure")

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
-
-TODO: model the propensity weights
-    - If I understand this right, we're modeling the propensity for a county to be located in a highly restrictive state.
-Is that right?
-So we want an estimated probability of a binary outcome, so we want logistic regression.
-Is that right?
 
@@
 library(stringr)
+library(kableExtra)
+library(table1)
 ```
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#load in data
+#load in relevant data
 providers <- read.csv("provider_list_geo.csv")
@@
 ########### calculate abortion rates in 2010 ###########
+
+#filter abortion rate data to only 2010
 abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
 
-# clean countydem data to make counties and states match abortion count data 
+# clean countydem data to make counties and states match those in abortion count data 
 countydem <- read.csv("county_demographics.csv")
@@
 
+abortions2010$county <- str_to_title(abortions2010$county)
+countydem$county <- str_to_title(countydem$county)
+
 abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
-abortions2010$county <- gsub(abortions2010$county, pattern = "ind. city", replacement = "city")
+abortions2010$county <- gsub(abortions2010$county, pattern = "Ind. City", replacement = "City")
 abortions2010$county <- gsub(abortions2010$county, pattern = "Saint", replacement = "St.")
-abortions2010$county <- gsub(abortions2010$county, pattern = "Dekalb", replacement = "DeKalb")
-abortions2010$county <- gsub(abortions2010$county, pattern = "Desoto", replacement = "DeSoto")
 abortions2010$county <- gsub(abortions2010$county, pattern = "Miami Dade", replacement = "Miami-Dade")
@@
 abortions2010$county <- gsub(abortions2010$county, pattern = "Kaua'i", replacement = "Kauai")
-abortions2010$county <- gsub(abortions2010$county, pattern = "City", replacement = "city")
-countydem$county <- gsub(countydem$county, pattern = "City", replacement = "city")
-abortions2010$county <- gsub(abortions2010$county, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
 abortions2010$county <- gsub(abortions2010$county, pattern = "St.e Genevieve", replacement = "Ste. Genevieve")
 abortions2010$county <- gsub(abortions2010$county, pattern = "Shannon/Oglala Lakota", replacement = "Oglala Lakota")
-abortions2010$county <- gsub(abortions2010$county, pattern = "Lamoure", replacement = "LaMoure")
 
+#Add number of women to abortion count data
 for (i in 1:nrow(abortions2010)) {
@@
 
-abortions2010 <- abortions2010 %>% filter(!(county %in% c("NONRESIDENTS", "PROTECTED", "OTHER STATES", "OTHER COUNTRIES", "NON-RESIDENTS", "OTHER", "NEVADA PROTECTED", "NEVADA UNSPECIFIED", "NEBRASKA PROTECTED", "county", "MARYLAND", "St.
-Louis County")))
+#remove non-counties
+abortions2010 <- abortions2010 %>% filter(!(county %in% c("Nonresidents", "Protected", "Other States", "Other Countries", "Non-Residents", "Other", "Nevada Protected", "Nevada Unspecified", "Nebraska Protected", "County", "Maryland")))
 
+#look at counties missing women pop data
 abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]
+#DeWitt, Bedford city, Clifton Forge city in abortion rate data  but not county dem data
 
-#make rates
+#calculate abortion rates
 abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)
+
+#Add abortion counts to county data
+for (i in 1:nrow(countydem)) {
+  if (length(abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]) == 0) {
+    countydem$AbortionRate2010[i] <- NA
+  }
+  else {
+    countydem$AbortionRate2010[i] <- abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]
+  }
+}
 
@@
 ```{r}
-## calculate percent democratic votes
+########### calculate percent democratic votes by county in in 2008 ###########
 countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
 countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)
-countypresdem2008$county_name[!(countypresdem2008$county_name %in% countydem$county)]
 
-countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "City", replacement = "city")
-countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Dekalb", replacement = "DeKalb")
-countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Desoto", replacement = "DeSoto")
-countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Lac Qui Parle", replacement = "Lac qui Parle")
+countydem$county <- gsub(countydem$county, pattern = " Parish", replacement = "")
+countydem$county <- gsub(countydem$county, pattern = " City", replacement = "")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " City", replacement = "")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = "Saint", replacement = "St.")
+countypresdem2008$county_name <- gsub(countypresdem2008$county_name, pattern = " County", replacement = "")
+
+missingcounties <- countypresdem2008[!(countypresdem2008$county_name %in% countydem$county), ]
 
 countypresdem2008 <- countypresdem2008 %>% mutate(percentdem2008 = candidatevotes/totalvotes) 
-names(countypresdem2008)
 countypresdem2008 <- countypresdem2008 %>% rename(state.abb = state_po, county = county_name) %>% dplyr::select(state.abb, county, percentdem2008)
@@
 ```{r}
-# calculate distances from county centroids to closest provider location
+########### calculate distances from county centroids to closest provider location ###########
 
@@
 ```{r}
-#make propensity scores
+#look at demographics
+
+countydem <- countydem %>% mutate(inanalysis = case_when(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI") ~ 1, !(state.abb %in% c("AZ", "DE", "GA", "IL", "IN", "MI", "NY", "NC", "OH", "OK", "OR", "PA", "SC", "TX", "UT", "VT", "WA", "WI")) ~ 0))
+
+countydemanalysis <- countydem %>% filter(inanalysis == 1)
+
+summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
+usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+summarytable <- left_join(summarytable, usedcounties, by = "state")
+summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
+names(countydem)
+countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+
+countydem$inanalysislabel <- factor(countydem$inanalysis, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))
+
+label(countydem$Women) = "Women"
+label(countydem$White.pct) = "White"
+label(countydem$Black.pct) = "Black"
+label(countydem$Native.American.pct) = "Native American"
+label(countydem$Asian.pct) = "Asian"
+label(countydem$Hispanic.pct) = "Hispanic"
+label(countydem$Median.Income) = "Median Income"
+label(countydem$College.Grad.pct) = "College Graduates"
+label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+
+units(countydem$Women) = "No."
+units(countydem$White.pct) = "%"
+units(countydem$Black.pct) = "%"
+units(countydem$Native.American.pct) = "%"
+units(countydem$Asian.pct) = "%"
+units(countydem$Hispanic.pct) = "%"
+units(countydem$Median.Income) = "$"
+units(countydem$College.Grad.pct) = "%"
+units(countydem$percentdem2008) = "%"
+
+table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, data = countydem)
+
+propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
+summary(propensitywts)
+
+p.score <- predict(propensitywts, countydem, type = "response")
+df <- countydem
+df$ps <- p.score
+
+#3 plots of distribution of propensity scores
+
+ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+
+ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
+  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+  facet_grid(as.factor(inanalysis) ~ .) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  scale_color_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical')
+
+df %>%
+  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+  group_by(inanalysis, ps.grp) %>%
+  summarize(n = n()) %>%
+  ungroup() %>%
+  mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
+  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
+  geom_bar(stat = 'identity', position = 'identity') +
+  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
+  xlab('Probability of Being in the Analysis') +
+  ylab('N') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
+  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+  theme(legend.position = 'bottom', legend.direction = 'vertical',
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank())
+
+```
+
+```{r}
+countydemanalysis <- df %>% filter(inanalysis == 1)
+countydemanalysis <- left_join(countydemanalysis, restrictions, by = "state")
+countycentroids <- countycentroids %>% rename(county = COUNAME, state = STNAME)
+countycentroids$county <- str_to_title(countycentroids$county)
+countycentroids <- countycentroids %>% select(county, state, distancetoprovider)
+countydemanalysis <- left_join(countydemanalysis, countycentroids, by = c("state", "county"))
+
+countydemanalysis$highorlow <- factor(countydemanalysis$highorlow, levels = c("low", "high"))
+model <- lm(AbortionRate2010 ~ highorlow + distancetoprovider + highorlow*distancetoprovider, data = countydemanalysis, weights = ps)
+summary(model)
 

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
+REPLICATION:
+TODO: make sure all the counties match/are represented <- Jo
+TODO: t tests for the demographic info for table 2 <- Toby
+TODO: averages of averages vs. averages for table 2 
+TODO: possibly email Alice asking about this county vs. state table 2 issue <- Jo 
+TODO: change variable name bc it's awful <- Jo
+TODO: write up <- Toby
+TODO: find abortion data used by Brown et al. to compare in discussion <- Toby
+
+EXTENSION:
+TODO: use all counties we have data on instead of just those 18 states
+TODO: rescore the high vs. low restriction variable to make it make more sense
+TODO: add in rural vs urban, % college grad, more variables do a quick variable selection
+TODO: repeat for mortality and fertility <- Jo look for this data
 

### 2021-12-14 paper_replication.Rmd
Added t-test p-values to Table 2

@@
 
-table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, data = countydem)
+# The following function is based on 
+# "Using the table1 Package to Create HTML Tables of Descriptive Statistics"
+# by Benjamin Rich.  Available at:
+# https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-a-column-of-p-values
+p_value_t_test <- function(x, ...){
+    y <- unlist(x[1:2])
+    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+}
 
+table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
+       data = countydem,
+       extra.col = list(`P-value` = p_value_t_test),
+       footnote = 'P-values are derived by two sample t-tests.'
+       )
+```
+
+```{r}
 propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ```{r}
-########### calculate abortion rates in 2010 ###########
-
-#filter abortion rate data to only 2010
-abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
-
-# clean countydem data to make counties and states match those in abortion count data 
+########### clean county dem data ###########
 countydem <- read.csv("county_demographics.csv")
@@
 
-abortions2010$county <- str_to_title(abortions2010$county)
 countydem$county <- str_to_title(countydem$county)
 
+```
+
+```{r}
+########### calculate abortion rates in 2010 ###########
+
+#filter abortion rate data to only 2010
+abortions2010 <- abortions %>% dplyr::select(state, county, X2010)
+
+length(unique(abortions2010$state)) #only 42 states in abortion data
+
+#look at percent missing abortion counts by state
+aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+
+# clean abortion data to make counties and states match county dem data
+abortions2010$county <- str_to_title(abortions2010$county)
 abortions2010$county <- gsub(abortions2010$county, pattern = "[()]", replacement = "")
@@
 
-#Add number of women to abortion count data
-for (i in 1:nrow(abortions2010)) {
-  if (length(countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]) == 0) {
-    abortions2010$Women[i] <- NA
-  }
-  else {
-    abortions2010$Women[i] <- countydem$Women[countydem$state.abb == abortions2010$state[i] & countydem$county == abortions2010$county[i]]
-  }
-}
-
-#fix NA values
+#fix NA values and make abortion counts numeric
 abortions2010$X2010 <- na_if(abortions2010$X2010, "?")
@@
 
-#look at counties missing women pop data
-abortions2010missingwomen <- abortions2010[is.na(abortions2010$Women), ]
-#DeWitt, Bedford city, Clifton Forge city in abortion rate data  but not county dem data
+#look at counties in abortion data but not countydem data
+abortions2010$county[!(abortions2010$county %in% countydem$county)]
+#"St. Louis County", "Bedford City", "Clifton Forge City"
 
-#calculate abortion rates
-abortions2010 <- abortions2010 %>% mutate(rate = X2010/Women)
+#look at counties in countydem data but not abortion data
+countydem$county[!(countydem$county %in% abortions2010$county)] #many 
+unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"
 
@@
 for (i in 1:nrow(countydem)) {
-  if (length(abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]) == 0) {
-    countydem$AbortionRate2010[i] <- NA
+  if (length(abortions2010$X2010[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]) == 0) {
+    countydem$AbortionCount2010[i] <- NA
   }
   else {
-    countydem$AbortionRate2010[i] <- abortions2010$rate[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]
+    countydem$AbortionCount2010[i] <- abortions2010$X2010[abortions2010$state == countydem$state.abb[i] & abortions2010$county == countydem$county[i]]
   }
 }
+
+#calculate abortion rates
+countydem <- countydem %>% mutate(AbortionRate2010 = AbortionCount2010/Women)
+
+#look at percent missing abortion rates by state
+aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
 
@@
 
-missingcounties <- countypresdem2008[!(countypresdem2008$county_name %in% countydem$county), ]
+missingcounties <- countypresdem2008[!(countypresdem2008$county_name %in% countydem$county), ] #all of AK and a few others (DC, Dona Ana New Mexico, Kansas City Missouri, Maine Uocava Maine, Federal Precinct Rhode Island)
 
@@
 countydem <- left_join(countydem, countypresdem2008, by = c("state.abb", "county"))
+
+#look at percent missing dem vote by state
+aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+
 ```
@@
 
+countycentroids <- countycentroids %>% rename(county = COUNAME, state = STNAME)
+countycentroids$county <- str_to_title(countycentroids$county)
+countycentroids <- countycentroids %>% select(county, state, distancetoprovider)
+countycentroids <- countycentroids %>% filter(state != "Puerto Rico")
+countycentroids$county <- gsub(countycentroids$county, pattern = " City", replacement = "")
+
+missingcounties <- countycentroids[!(countycentroids$county %in% countydem$county), ] #Alaska and DC
+
+countydem <- left_join(countydem, countycentroids, by = c("state", "county"))
+
+#look at percent missing distance by state
+aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+
 ```
@@
 restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
-                                        highorlow = case_when(totalrestrict >= 3 ~ "high", 
+                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                               totalrestrict < 3 ~ "low"))
+restrictions$state[!(restrictions$state %in% countydem$state)] #none
+countydem <- left_join(countydem, restrictions, by = "state")
+countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))
+
 ```
@@
 ```{r}
+#calculate propensity scores
 propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
@@
 ```{r}
-countydemanalysis <- df %>% filter(inanalysis == 1)
-countydemanalysis <- left_join(countydemanalysis, restrictions, by = "state")
-countycentroids <- countycentroids %>% rename(county = COUNAME, state = STNAME)
-countycentroids$county <- str_to_title(countycentroids$county)
-countycentroids <- countycentroids %>% select(county, state, distancetoprovider)
-countydemanalysis <- left_join(countydemanalysis, countycentroids, by = c("state", "county"))
+#run models
+countydemanalysis <- countydem %>% filter(inanalysis == 1)
 
-countydemanalysis$highorlow <- factor(countydemanalysis$highorlow, levels = c("low", "high"))
-model <- lm(AbortionRate2010 ~ highorlow + distancetoprovider + highorlow*distancetoprovider, data = countydemanalysis, weights = ps)
+model <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
 summary(model)

### 2021-12-14 paper_replication.Rmd
Update paper_replication.Rmd

@@
 REPLICATION:
-TODO: make sure all the counties match/are represented <- Jo
 TODO: t tests for the demographic info for table 2 <- Toby
@@
 TODO: possibly email Alice asking about this county vs. state table 2 issue <- Jo 
-TODO: change variable name bc it's awful <- Jo
 TODO: write up <- Toby

### 2021-12-15 report.Rmd
Partial draft of text for the write up.

---
title: "Report"
output: pdf_document
---

## Introduction

In this paper we replicate an analysis published in JAMA Network Open by Benjamin P. Brown, MD, MS; Luciana E. Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  We then extend this analysis by considering . . .

The data set from the Brown et al. article was not made public, and due to limitations of publicly available data, we have narrowed the scope of our replication to analyzing a single year. Specifically we analyzed data from 2010, primarily because this was the year within the original paper's time range for which we had the most detailed information on state legislation restricting abortions.

## Study Design

We used a propensity weighted linear model to assess the association of restrictive abortion legislation and abortion rates.  We used data from the same set of 18 states as Brown et al., and used propensity weighting to account for differences between those states and the other 32 states.  The propensity weights were calculated using the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008.

## Participants, Variables and Data Sources

We used abortion count data compiled by Wm. Robert Johnston and available at www.johnstonsarchive.net/policy/abortion.  At least partial data was available for all 18 of the states included in the analysis by Brown et al., and for most of those states we had data for more than 80% of the counties (see Table 1).  The percentages of counties reporting were similar to the percentages reporting in Brown et al.

For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, targeted restrictions on abortion providers (TRAP) laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors. Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3/4 types of laws.

The covariates for the propensity score model were taken from the US Census Bureau

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 REPLICATION:
-TODO: t tests for the demographic info for table 2 <- Toby
 TODO: averages of averages vs. averages for table 2 
@@
 
+```{r}
+natality <- read.delim("natalitydata.txt")
+maternalmortality <- read.delim("maternalmortality.txt")
+infantdeath <- read.delim("infantdeathrecords.txt")
+
+infantdeath$County <- gsub(infantdeath$County, pattern = ", [[:print:]]*$", replacement = "")
+infantdeath$County <- gsub(infantdeath$County, pattern = " County", replacement = "")
+infantdeath$County <- str_to_title(infantdeath$County)
+infantdeath$County <- na_if(infantdeath$County, "")
+infantdeath <- infantdeath %>% filter(County != "Unidentified Counties")
+infantdeath$County <- gsub(infantdeath$County, pattern = " Parish", replacement = "")
+infantdeath$County <- gsub(infantdeath$County, pattern = "Borough", replacement = "Municipality")
+infantdeath$County <- gsub(infantdeath$County, pattern = " City", replacement = "")
+infantdeath$County[!(infantdeath$County %in% countydem$county)]
+infantdeath <- infantdeath %>% rename(state = State, county = County, infantdeathrate = Death.Rate)
+infantdeath <- infantdeath %>% dplyr::select(state, county, infantdeathrate)
+
+maternalmortality$County <- gsub(maternalmortality$County, pattern = ", [[:print:]]*$", replacement = "")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = " County", replacement = "")
+maternalmortality$County <- str_to_title(maternalmortality$County)
+maternalmortality$County <- na_if(maternalmortality$County, "")
+maternalmortality <- maternalmortality %>% filter(County != "Unidentified Counties")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = " Parish", replacement = "")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = "Borough", replacement = "Municipality")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = " City", replacement = "")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = "La Porte", replacement = "Laporte")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = "De Kalb", replacement = "Dekalb")
+maternalmortality$County <- gsub(maternalmortality$County, pattern = "Mc Kean", replacement = "Mckean")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "Municipality", replacement = "Borough")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "Borough", replacement = "And Borough")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = "/", replacement = " ")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = " Census Area", replacement = "")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = " And", replacement = "")
+maternalmortality$County[!(maternalmortality$County %in% countydem$county)]
+
+maternalmortality <- maternalmortality %>% rename(state = State, county = County, maternaldeaths = Deaths, maternaldeathrate = Crude.Rate)
+maternalmortality <- maternalmortality %>% dplyr::select(state, county, maternaldeathrate)
+
+natality$County <- gsub(natality$County, pattern = ", [[:print:]]*$", replacement = "")
+natality$County <- gsub(natality$County, pattern = " County", replacement = "")
+natality$County <- str_to_title(natality$County)
+natality$County <- na_if(natality$County, "")
+natality <- natality %>% filter(County != "Unidentified Counties")
+natality$County <- gsub(natality$County, pattern = " Parish", replacement = "")
+natality$County <- gsub(natality$County, pattern = "Borough", replacement = "Municipality")
+natality$County <- gsub(natality$County, pattern = " City", replacement = "")
+natality$County <- gsub(natality$County, pattern = "La Porte", replacement = "Laporte")
+natality$County[!(natality$County %in% countydem$county)]
+natality <- natality %>% rename(state = State, county = County)
+natality <- natality %>% dplyr::select(state, county, Birth.Rate, Fertility.Rate)
+
+countydem <- left_join(countydem, infantdeath, by = c("state", "county"))
+countydem <- left_join(countydem, maternalmortality, by = c("state", "county"))
+countydem <- left_join(countydem, natality, by = c("state", "county"))
+
+sum(is.na(countydem$infantdeathrate)/nrow(countydem))
+countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
+sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
+sum(is.na(countydem$Birth.Rate)/nrow(countydem))
+
+```
+
 \newpage

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 countycentroids$county <- str_to_title(countycentroids$county)
-countycentroids <- countycentroids %>% select(county, state, distancetoprovider)
+countycentroids <- countycentroids %>% dplyr::select(county, state, distancetoprovider)
 countycentroids <- countycentroids %>% filter(state != "Puerto Rico")

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 #run models
-countydemanalysis <- countydem %>% filter(inanalysis == 1)
+countydemanalysis <- df %>% filter(inanalysis == 1)
 
@@
 ```{r}
+####### add in natality, maternalmortality, and infantdeath data #######
 natality <- read.delim("natalitydata.txt")
@@
 countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
+countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Missing")
+
 sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
 sum(is.na(countydem$Birth.Rate)/nrow(countydem))
+
+```
+
+```{r}
+#### recode restriction level and include all counties with abortion data in analysis ####
+
+countydem <- countydem %>% mutate(hasabortiondata = case_when(is.na(AbortionRate2010) ~ 0, 
+                                                              !(is.na(AbortionRate2010)) ~ 1))
+
+countydemanalysis <- countydem %>% filter(hasabortiondata == 1)
+
+summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+summarytable <- left_join(summarytable, usedcounties, by = "state")
+summarytable
+summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
+names(countydem)
+countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+
+countydem$hasabortiondatalabel <- factor(countydem$hasabortiondata, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))
+
+label(countydem$Women) = "Women"
+label(countydem$White.pct) = "White"
+label(countydem$Black.pct) = "Black"
+label(countydem$Native.American.pct) = "Native American"
+label(countydem$Asian.pct) = "Asian"
+label(countydem$Hispanic.pct) = "Hispanic"
+label(countydem$Median.Income) = "Median Income"
+label(countydem$College.Grad.pct) = "College Graduates"
+label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+
+units(countydem$Women) = "No."
+units(countydem$White.pct) = "%"
+units(countydem$Black.pct) = "%"
+units(countydem$Native.American.pct) = "%"
+units(countydem$Asian.pct) = "%"
+units(countydem$Hispanic.pct) = "%"
+units(countydem$Median.Income) = "$"
+units(countydem$College.Grad.pct) = "%"
+units(countydem$percentdem2008) = "%"
+
+# The following function is based on 
+# "Using the table1 Package to Create HTML Tables of Descriptive Statistics"
+# by Benjamin Rich.  Available at:
+# https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-a-column-of-p-values
+p_value_t_test <- function(x, ...){
+    y <- unlist(x[1:2])
+    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+}
+
+table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasabortiondatalabel, 
+       data = countydem,
+       extra.col = list(`P-value` = p_value_t_test),
+       footnote = 'P-values are derived by two sample t-tests.'
+       )
+
+#calculate propensity scores
+#include education this time
+propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
+summary(propensitywts)
+
+p.score <- predict(propensitywts, countydem, type = "response")
+df <- countydem
+df$ps <- p.score
+
+#3 plots of distribution of propensity scores
+
+ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+
+ggplot(df, aes(x = ps, fill = hasabortiondatalabel, color = hasabortiondatalabel)) +
+  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+  facet_grid(as.factor(hasabortiondata) ~ .) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  scale_color_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical')
+
+df %>%
+  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+  group_by(hasabortiondata, ps.grp) %>%
+  summarize(n = n()) %>%
+  ungroup() %>%
+  mutate(n2 = ifelse(hasabortiondata == 0, yes = n, no =  -1*n)) %>%
+  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasabortiondata))) +
+  geom_bar(stat = 'identity', position = 'identity') +
+  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasabortiondata == 0, 8, -8))) +
+  xlab('Probability of Being in the Analysis') +
+  ylab('N') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
+  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+  theme(legend.position = 'bottom', legend.direction = 'vertical',
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank())
+
+#run model
+countydemanalysis <- df %>% filter(hasabortiondata == 1)
+
+model <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
+summary(model)
+
+```
+
+```{r}
+###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
+countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
+countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
+countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
+hist(countydem$maternaldeathrate)
+
+countydem$infantdeathrate <- na_if(countydem$infantdeathrate, "Missing County")
+countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "[()]", replacement = "")
+countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "Unreliable", replacement = "")
+countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
+hist(countydem$infantdeathrate)
+
+countydem$Birth.Rate <- na_if(countydem$Birth.Rate, "Missing County")
+countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
+hist(countydem$Birth.Rate)
+
+countydem$Fertility.Rate <- na_if(countydem$Fertility.Rate, "Missing County")
+countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
+hist(countydem$Fertility.Rate)
+
+```
+```{r}
+### repeat analysis for infant death rate ###
+countydem <- countydem %>% mutate(hasinfantdeathdata = case_when(is.na(countydem$infantdeathrate) ~ 0, 
+                                                              !(is.na(countydem$infantdeathrate)) ~ 1))
+
+countydeminfantdeathanalysis <- countydem %>% filter(hasinfantdeathdata == 1)
+
+summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+usedcounties <- aggregate(infantdeathrate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+summarytable <- left_join(summarytable, usedcounties, by = "state")
+summarytable %>% mutate(percent = infantdeathrate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Data Included in Analysis by State") %>% kable_styling()
+names(countydem)
+countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+
+countydem$hasinfantdeathdatalabel <- factor(countydem$hasinfantdeathdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
+
+label(countydem$Women) = "Women"
+label(countydem$White.pct) = "White"
+label(countydem$Black.pct) = "Black"
+label(countydem$Native.American.pct) = "Native American"
+label(countydem$Asian.pct) = "Asian"
+label(countydem$Hispanic.pct) = "Hispanic"
+label(countydem$Median.Income) = "Median Income"
+label(countydem$College.Grad.pct) = "College Graduates"
+label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+
+units(countydem$Women) = "No."
+units(countydem$White.pct) = "%"
+units(countydem$Black.pct) = "%"
+units(countydem$Native.American.pct) = "%"
+units(countydem$Asian.pct) = "%"
+units(countydem$Hispanic.pct) = "%"
+units(countydem$Median.Income) = "$"
+units(countydem$College.Grad.pct) = "%"
+units(countydem$percentdem2008) = "%"
+
+p_value_t_test <- function(x, ...){
+    y <- unlist(x[1:2])
+    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+}
+
+print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasinfantdeathdatalabel, 
+       data = countydem,
+       extra.col = list(`P-value` = p_value_t_test),
+       footnote = 'P-values are derived by two sample t-tests.'
+       ))
+
+#calculate propensity scores
+#include education this time
+propensitywts <- glm(hasinfantdeathdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
+
+p.score <- predict(propensitywts, countydem, type = "response")
+df <- countydem
+df$ps <- p.score
+
+#3 plots of distribution of propensity scores
+
+plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+print(plot1)
+
+plot2 <- ggplot(df, aes(x = ps, fill = hasinfantdeathdatalabel, color = hasinfantdeathdatalabel)) +
+  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+  facet_grid(as.factor(hasinfantdeathdata) ~ .) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  scale_color_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical')
+print(plot2)
+
+plot3 <- df %>%
+  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+  group_by(hasinfantdeathdata, ps.grp) %>%
+  summarize(n = n()) %>%
+  ungroup() %>%
+  mutate(n2 = ifelse(hasinfantdeathdata == 0, yes = n, no =  -1*n)) %>%
+  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasinfantdeathdata))) +
+  geom_bar(stat = 'identity', position = 'identity') +
+  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasinfantdeathdata == 0, 8, -8))) +
+  xlab('Probability of Being in the Analysis') +
+  ylab('N') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
+  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+  theme(legend.position = 'bottom', legend.direction = 'vertical',
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank())
+print(plot3)
+
+#run model
+countydeminfantdeathanalysis <- df %>% filter(hasinfantdeathdata == 1)
+
+model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
+summary(model)
+
+```
+
+```{r}
+### repeat analysis for birth rate ###
+
+countydem <- countydem %>% mutate(hasbirthratedata = case_when(is.na(countydem$Birth.Rate) ~ 0, 
+                                                              !(is.na(countydem$Birth.Rate)) ~ 1))
+
+countydembirthrateanalysis <- countydem %>% filter(hasbirthratedata == 1)
+
+summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+usedcounties <- aggregate(Birth.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+usedcounties
+summarytable <- left_join(summarytable, usedcounties, by = "state")
+summarytable
+summarytable %>% mutate(percent = Birth.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Data Included in Analysis by State") %>% kable_styling()
+names(countydem)
+countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+
+countydem$hasbirthratedatalabel <- factor(countydem$hasbirthratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
+
+label(countydem$Women) = "Women"
+label(countydem$White.pct) = "White"
+label(countydem$Black.pct) = "Black"
+label(countydem$Native.American.pct) = "Native American"
+label(countydem$Asian.pct) = "Asian"
+label(countydem$Hispanic.pct) = "Hispanic"
+label(countydem$Median.Income) = "Median Income"
+label(countydem$College.Grad.pct) = "College Graduates"
+label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+
+units(countydem$Women) = "No."
+units(countydem$White.pct) = "%"
+units(countydem$Black.pct) = "%"
+units(countydem$Native.American.pct) = "%"
+units(countydem$Asian.pct) = "%"
+units(countydem$Hispanic.pct) = "%"
+units(countydem$Median.Income) = "$"
+units(countydem$College.Grad.pct) = "%"
+units(countydem$percentdem2008) = "%"
+
+p_value_t_test <- function(x, ...){
+    y <- unlist(x[1:2])
+    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+}
+
+print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasbirthratedatalabel, 
+       data = countydem,
+       extra.col = list(`P-value` = p_value_t_test),
+       footnote = 'P-values are derived by two sample t-tests.'
+       ))
+
+#calculate propensity scores
+#include education this time
+propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
+summary(propensitywts)
+
+p.score <- predict(propensitywts, countydem, type = "response")
+df <- countydem
+df$ps <- p.score
+#3 plots of distribution of propensity scores
+
+plot1 <- ggplot(df, aes(x = ps, group = hasbirthratedatalabel)) + geom_density(aes(fill = hasbirthratedatalabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+print(plot1)
+
+plot2 <- ggplot(df, aes(x = ps, fill = hasbirthratedatalabel, color = hasbirthratedatalabel)) +
+  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+  facet_grid(as.factor(hasbirthratedata) ~ .) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  scale_color_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical')
+print(plot2)
+
+plot3 <- df %>%
+  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+  group_by(hasbirthratedata, ps.grp) %>%
+  summarize(n = n()) %>%
+  ungroup() %>%
+  mutate(n2 = ifelse(hasbirthratedata == 0, yes = n, no =  -1*n)) %>%
+  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasbirthratedata))) +
+  geom_bar(stat = 'identity', position = 'identity') +
+  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasbirthratedata == 0, 8, -8))) +
+  xlab('Probability of Being in the Analysis') +
+  ylab('N') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
+  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+  theme(legend.position = 'bottom', legend.direction = 'vertical',
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank())
+print(plot3)
+
+#run model
+countydembirthrateanalysis <- df %>% filter(hasbirthratedata == 1)
+
+model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
+summary(model)
+
+test <- countydembirthrateanalysis %>% filter(ps > 0.9999999)
+```
+
+```{r}
+### repeat analysis for fertility rate ###
+
+countydem <- countydem %>% mutate(hasfertilityratedata = case_when(is.na(countydem$Fertility.Rate) ~ 0, 
+                                                              !(is.na(countydem$Fertility.Rate)) ~ 1))
+
+countydemfertilityrateanalysis <- countydem %>% filter(hasfertilityratedata == 1)
+
+summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+usedcounties <- aggregate(Fertility.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+usedcounties
+summarytable <- left_join(summarytable, usedcounties, by = "state")
+summarytable
+summarytable %>% mutate(percent = Fertility.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Data Included in Analysis by State") %>% kable_styling()
+names(countydem)
+countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+
+countydem$hasfertilityratedatalabel <- factor(countydem$hasfertilityratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
+
+label(countydem$Women) = "Women"
+label(countydem$White.pct) = "White"
+label(countydem$Black.pct) = "Black"
+label(countydem$Native.American.pct) = "Native American"
+label(countydem$Asian.pct) = "Asian"
+label(countydem$Hispanic.pct) = "Hispanic"
+label(countydem$Median.Income) = "Median Income"
+label(countydem$College.Grad.pct) = "College Graduates"
+label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+
+units(countydem$Women) = "No."
+units(countydem$White.pct) = "%"
+units(countydem$Black.pct) = "%"
+units(countydem$Native.American.pct) = "%"
+units(countydem$Asian.pct) = "%"
+units(countydem$Hispanic.pct) = "%"
+units(countydem$Median.Income) = "$"
+units(countydem$College.Grad.pct) = "%"
+units(countydem$percentdem2008) = "%"
+
+p_value_t_test <- function(x, ...){
+    y <- unlist(x[1:2])
+    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+}
+
+print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasfertilityratedatalabel, 
+       data = countydem,
+       extra.col = list(`P-value` = p_value_t_test),
+       footnote = 'P-values are derived by two sample t-tests.'
+       ))
+
+#calculate propensity scores
+#include education this time
+propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
+summary(propensitywts)
+
+p.score <- predict(propensitywts, countydem, type = "response")
+df <- countydem
+df$ps <- p.score
+#3 plots of distribution of propensity scores
+
+plot1 <- ggplot(df, aes(x = ps, group = hasfertilityratedatalabel)) + geom_density(aes(fill = hasfertilityratedatalabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+print(plot1)
+
+plot2 <- ggplot(df, aes(x = ps, fill = hasfertilityratedatalabel, color = hasfertilityratedatalabel)) +
+  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+  facet_grid(as.factor(hasfertilityratedata) ~ .) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  scale_color_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical')
+print(plot2)
+
+plot3 <- df %>%
+  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+  group_by(hasfertilityratedata, ps.grp) %>%
+  summarize(n = n()) %>%
+  ungroup() %>%
+  mutate(n2 = ifelse(hasfertilityratedata == 0, yes = n, no =  -1*n)) %>%
+  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasfertilityratedata))) +
+  geom_bar(stat = 'identity', position = 'identity') +
+  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasfertilityratedata == 0, 8, -8))) +
+  xlab('Probability of Being in the Analysis') +
+  ylab('N') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
+  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+  theme(legend.position = 'bottom', legend.direction = 'vertical',
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank())
+print(plot3)
+
+#run model
+countydemfertilityrateanalysis <- df %>% filter(hasfertilityratedata == 1)
+
+model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
+summary(model)
 

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 EXTENSION:
-TODO: use all counties we have data on instead of just those 18 states
-TODO: rescore the high vs. low restriction variable to make it make more sense
-TODO: add in rural vs urban, % college grad, more variables do a quick variable selection
-TODO: repeat for mortality and fertility <- Jo look for this data
+TODO: write literature review and describe motivation for extension
+TODO: figure out what is happening with the propensity scores for the extension analyses
+TODO: decide whether to rescore the high vs. low restriction variable as a factor or an integer (currently coded as the latter)
+TODO: add in rural vs urban and more variables do a quick variable selection
 

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 TODO: averages of averages vs. averages for table 2
-TODO: possibly email Alice asking about this county vs. state table 2 issue <- Jo
-TODO: write up <- Toby
+TODO: the paper says "all models included state and year fixed effects and standard errors adjusted for state-level clustering" -> does this mean we need to include states as a variable in our model?
+TODO: the paper says "We addressed data gaps in demographic variables by using the estimate from the nearest year for which an estimate was available", so we need to include this methodology
+TODO: bias analysis as done in the paper, comparing Guttmacher Census State Data to our State Data
+TODO: run the models without the distance variable as well as with the distance variable as the outcome, as done in the paper
+TODO: write up everything: we need to follow the format of the original paper <- Toby
 TODO: find abortion data used by Brown et al. to compare in discussion <- Toby
@@
 TODO: decide whether to rescore the high vs. low restriction variable as a factor or an integer (currently coded as the latter)
-TODO: add in rural vs urban and more variables do a quick variable selection
+TODO: add in rural vs urban variable
+TODO: look at Hawkins et al paper linked in the original paper to see what mortality data they used. consider using mortality data by state.
 

### 2021-12-15 paper_replication.Rmd
TODO list update

@@
 TODO: averages of averages vs. averages for table 2
+
 TODO: the paper says "all models included state and year fixed effects and standard errors adjusted for state-level clustering" -> does this mean we need to include states as a variable in our model?
+-- No, they did that because they were doing the longitudinal analysis, so like abortion rates in a given state for one year are not independent of the next, so they're accounting for that.
+
 TODO: the paper says "We addressed data gaps in demographic variables by using the estimate from the nearest year for which an estimate was available", so we need to include this methodology

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
   xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
+  ggtitle('Figure 1: Propensity Score Distribution by Group') +
   scale_fill_discrete('') +
@@
 
-ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
-  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-  facet_grid(as.factor(inanalysis) ~ .) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  scale_color_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical')
-
-df %>%
-  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-  group_by(inanalysis, ps.grp) %>%
-  summarize(n = n()) %>%
-  ungroup() %>%
-  mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
-  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
-  geom_bar(stat = 'identity', position = 'identity') +
-  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
-  xlab('Probability of Being in the Analysis') +
-  ylab('N') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
-  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-  theme(legend.position = 'bottom', legend.direction = 'vertical',
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank())
+# ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
+#   geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
+#   facet_grid(as.factor(inanalysis) ~ .) +
+#   xlab('Probability of Being in the Analysis') +
+#   ggtitle('Propensity Score Distribution by Group') +
+#   scale_fill_discrete('') +
+#   scale_color_discrete('') +
+#   theme(legend.position = 'bottom', legend.direction = 'vertical')
+# 
+# df %>%
+#   mutate(ps.grp = round(ps/0.05) * 0.05) %>%
+#   group_by(inanalysis, ps.grp) %>%
+#   summarize(n = n()) %>%
+#   ungroup() %>%
+#   mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
+#   ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
+#   geom_bar(stat = 'identity', position = 'identity') +
+#   geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
+#   xlab('Probability of Being in the Analysis') +
+#   ylab('N') +
+#   ggtitle('Propensity Score Distribution by Group') +
+#   scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
+#   scale_x_continuous(breaks = seq(0, 1, 0.05)) +
+#   theme(legend.position = 'bottom', legend.direction = 'vertical',
+#         axis.ticks.y = element_blank(),
+#         axis.text.y = element_blank())
 

### 2021-12-15 paper_replication.Rmd
Added linear model summaries (tables 3 and 4)

@@
 
-model <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
-summary(model)
+model_no_dist <- lm(AbortionRate2010 ~ restrictionlevel, data = countydemanalysis, weights = ps)
+summ_no_dist <- summary(model_no_dist)
+
+model_dist <- lm(AbortionRate2010 ~ restrictionlevel + distancetoprovider + restrictionlevel*distancetoprovider, data = countydemanalysis, weights = ps)
+summ_dist <- summary(model_dist)
+
+# Model Summaries (Note coefficients are per 1000 women)
+table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2]*1000,
+                  "p-value" = c('< 0.001')) %>% 
+  kbl(caption = "Table 3: Change in Abortions per 1000 Women in Model without Distances", 
+          digits = 3)
+
+p <- summ_dist$coefficients[4,4] %>% round(3)
+table_4 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
+                           model_dist$coefficients[4]*1000),
+                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
+  kbl(caption = "Table 4: Change in Abortions per 1000 Women in Model with Distances")
+print(table_4)
 

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
           digits = 3)
+print(table_3)
 

### 2021-12-15 report.Rmd
Update report.Rmd

@@
 Brown, MD, MS; Luciana E.
-Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  We then extend this analysis by considering .
-.
-.
+Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  Brown et al.
+use a binary classification of states legislative climates on abortion as either "highly restrictive" or not, and they assess the association with abortion rates using propensity weighted linear difference in differences models, based on data from 18 states.
 
@@
 
+We first reproduce the analysis of Brown et al.
+using data from the same 18 states.
+We then expand the analysis by using data from all 50 states and using an integer variable for the extent of abortion restriction legislation.
+Finally we extend the analysis to include associations of restrictive abortion legislation on  maternity mortality data, infant mortality data, and birth rate.
+
 ## Study Design
 
-We used a propensity weighted linear model to assess the association of restrictive abortion legislation and abortion rates.
-We used data from the same set of 18 states as Brown et al., and used propensity weighting to account for differences between those states and the other 32 states.
+We used propensity weighted linear models to assess the association of restrictive abortion legislation and abortion rates.
+We used data from the same set of 18 states as Brown et al.
+(see Table 1), and used propensity weighting to account for differences between those states and the other 32 states.
 The propensity weights were calculated using the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008.
+We fit linear models with and without including distances to abortion providers and an interaction term between those distances and restrictive legislation.
 
@@
 Robert Johnston and available at www.johnstonsarchive.net/policy/abortion.
-At least partial data was available for all 18 of the states included in the analysis by Brown et al., and for most of those states we had data for more than 80% of the counties (see Table 1).
+Abortion rates were calculated as number of abortions per female population.
+At least partial data was available for all 18 of the states included in the analysis by Brown et al., and for most of those states there was data for more than 80% of the counties (see Table 1).
 The percentages of counties reporting were similar to the percentages reporting in Brown et al.
+(see their Table 1).
 
@@
 
-The covariates for the propensity score model were taken from the US Census Bureau
+We estimated distances to providers as the straight-line distance from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
+
+## Results
+
+The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).
+This suggest that the 18 states included in the analysis are not exceptionally different from those not included regarding the demographic covariates, and accordingly we expect the propensity weights to accurately adjust for the differences.
+
+Tables 3 and 4 show the main results from our replication.
+Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a  reduction in abortions of 0.922 per 1,000 women, and that effect was reduced to 0.828 when adjusting for distances to providers.
+These effects were roughly twice as large as the effects calculated by Brown et al., nevertheless our estimates were within their reported 95% confidence intervals (see Brown et al.
+Table 3).
+
+## Limitations
+
+The original analysis and our replication are limited by the use of data from only 18 states.
+We address this shortcoming in our extension below.
+The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.

### 2021-12-15 paper_replication.Rmd
Added result tables integer restrictions, 50 state

@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ########### clean county dem data ###########
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ########### calculate abortion rates in 2010 ###########
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ########### calculate percent democratic votes by county in in 2008 ###########
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ########### calculate distances from county centroids to closest provider location ###########
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #make new variable for high vs. low restrictions
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #look at demographics
@@
 Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
-names(countydem)
+# names(countydem)
 countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #calculate propensity scores
@@
 
-# ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
-#   geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-#   facet_grid(as.factor(inanalysis) ~ .) +
-#   xlab('Probability of Being in the Analysis') +
-#   ggtitle('Propensity Score Distribution by Group') +
-#   scale_fill_discrete('') +
-#   scale_color_discrete('') +
-#   theme(legend.position = 'bottom', legend.direction = 'vertical')
-# 
-# df %>%
-#   mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-#   group_by(inanalysis, ps.grp) %>%
-#   summarize(n = n()) %>%
-#   ungroup() %>%
-#   mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
-#   ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
-#   geom_bar(stat = 'identity', position = 'identity') +
-#   geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
-#   xlab('Probability of Being in the Analysis') +
-#   ylab('N') +
-#   ggtitle('Propensity Score Distribution by Group') +
-#   scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
-#   scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-#   theme(legend.position = 'bottom', legend.direction = 'vertical',
-#         axis.ticks.y = element_blank(),
-#         axis.text.y = element_blank())
-
 ```
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #run models
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ####### add in natality, maternalmortality, and infantdeath data #######
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #### recode restriction level and include all counties with abortion data in analysis ####
@@
 Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
-names(countydem)
+# names(countydem)
 countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
@@
 
-#run model
+#run models, all 50 states, integer restrictions
 countydemanalysis <- df %>% filter(hasabortiondata == 1)
 
-model <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
-summary(model)
+ext_model_dist <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
+summ_ext_dist <- summary(ext_model_dist)
+
+ext_model_no_dist <- lm(AbortionRate2010 ~ totalrestrict, data = countydemanalysis, weights = ps)
+summ_ext_no_dist <- summary(ext_model_no_dist)
+
+# Model Summaries (Note coefficients are per 1000 women)
+table_ext_no_dist <- tibble(Characteristic = "Number of Types of Restrictive Legislation",
+                  Beta = ext_model_no_dist$coefficients[2]*1000,
+                  "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
+  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model without Distances, all 50 States",
+          digits = 3)
+print(table_ext_no_dist)
+
+table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
+                                            "Distance to Provider", 
+                                            "Distance/Restriction Interaction"),
+                  Beta = ext_model_dist$coefficients[2:4]*1000,
+                  "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
+  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
+print(table_ext_dist)
 
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
@@
 ```
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ### repeat analysis for infant death rate ###
@@
 Number of Counties with Data Included in Analysis by State") %>% kable_styling()
-names(countydem)
+# names(countydem)
 countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
@@
 
-plot2 <- ggplot(df, aes(x = ps, fill = hasinfantdeathdatalabel, color = hasinfantdeathdatalabel)) +
-  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-  facet_grid(as.factor(hasinfantdeathdata) ~ .) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  scale_color_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical')
-print(plot2)
-
-plot3 <- df %>%
-  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-  group_by(hasinfantdeathdata, ps.grp) %>%
-  summarize(n = n()) %>%
-  ungroup() %>%
-  mutate(n2 = ifelse(hasinfantdeathdata == 0, yes = n, no =  -1*n)) %>%
-  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasinfantdeathdata))) +
-  geom_bar(stat = 'identity', position = 'identity') +
-  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasinfantdeathdata == 0, 8, -8))) +
-  xlab('Probability of Being in the Analysis') +
-  ylab('N') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
-  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-  theme(legend.position = 'bottom', legend.direction = 'vertical',
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank())
-print(plot3)
-
 #run model
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ### repeat analysis for birth rate ###
@@
 
-plot2 <- ggplot(df, aes(x = ps, fill = hasbirthratedatalabel, color = hasbirthratedatalabel)) +
-  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-  facet_grid(as.factor(hasbirthratedata) ~ .) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  scale_color_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical')
-print(plot2)
-
-plot3 <- df %>%
-  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-  group_by(hasbirthratedata, ps.grp) %>%
-  summarize(n = n()) %>%
-  ungroup() %>%
-  mutate(n2 = ifelse(hasbirthratedata == 0, yes = n, no =  -1*n)) %>%
-  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasbirthratedata))) +
-  geom_bar(stat = 'identity', position = 'identity') +
-  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasbirthratedata == 0, 8, -8))) +
-  xlab('Probability of Being in the Analysis') +
-  ylab('N') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
-  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-  theme(legend.position = 'bottom', legend.direction = 'vertical',
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank())
-print(plot3)
-
 #run model
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ### repeat analysis for fertility rate ###
@@
 
-plot2 <- ggplot(df, aes(x = ps, fill = hasfertilityratedatalabel, color = hasfertilityratedatalabel)) +
-  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-  facet_grid(as.factor(hasfertilityratedata) ~ .) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  scale_color_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical')
-print(plot2)
-
-plot3 <- df %>%
-  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-  group_by(hasfertilityratedata, ps.grp) %>%
-  summarize(n = n()) %>%
-  ungroup() %>%
-  mutate(n2 = ifelse(hasfertilityratedata == 0, yes = n, no =  -1*n)) %>%
-  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasfertilityratedata))) +
-  geom_bar(stat = 'identity', position = 'identity') +
-  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasfertilityratedata == 0, 8, -8))) +
-  xlab('Probability of Being in the Analysis') +
-  ylab('N') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
-  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-  theme(legend.position = 'bottom', legend.direction = 'vertical',
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank())
-print(plot3)
-
 #run model

### 2021-12-15 paper_replication.Rmd
Formatting tables and figures

@@
 
-length(unique(abortions2010$state)) #only 42 states in abortion data
+# length(unique(abortions2010$state)) #only 42 states in abortion data
 
 #look at percent missing abortion counts by state
-aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+# aggregate(X2010 ~ state, data = abortions2010, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
 
@@
 #look at counties in abortion data but not countydem data
-abortions2010$county[!(abortions2010$county %in% countydem$county)]
+# abortions2010$county[!(abortions2010$county %in% countydem$county)]
 #"St. Louis County", "Bedford City", "Clifton Forge City"
@@
 #look at counties in countydem data but not abortion data
-countydem$county[!(countydem$county %in% abortions2010$county)] #many 
-unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"
+# countydem$county[!(countydem$county %in% abortions2010$county)] #many 
+# unique(countydem$state.abb[!(countydem$state.abb %in% abortions2010$state)]) #"AK" "CT" "IA" "LA" "MA" "NH" "NM" "WY"
 
@@
 #look at percent missing abortion rates by state
-aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+# aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
 
@@
 #look at percent missing dem vote by state
-aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+# aggregate(percentdem2008 ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
 
@@
 #look at percent missing distance by state
-aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
+# aggregate(distancetoprovider ~ state, data = countydem, function(x) {sum(!(is.na(x)))/length(x)}, na.action = NULL)
 
@@
                                                               totalrestrict < 3 ~ "low"))
-restrictions$state[!(restrictions$state %in% countydem$state)] #none
+# restrictions$state[!(restrictions$state %in% countydem$state)] #none
 countydem <- left_join(countydem, restrictions, by = "state")
@@
 
-summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
+table_1 <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
 usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-summarytable <- left_join(summarytable, usedcounties, by = "state")
-summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
+table_1 <- left_join(table_1, usedcounties, by = "state")
+table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling()
+
+to_adjust_kable_table_numbers <- kable(table_1)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # names(countydem)
@@
        extra.col = list(`P-value` = p_value_t_test),
-       footnote = 'P-values are derived by two sample t-tests.'
+       footnote = 'P-values are derived by two sample t-tests.',
+       caption = "County Data for States Included and Not Included in Analysis"
        )
@@
 propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
-summary(propensitywts)
+# summary(propensitywts)
 
@@
 
-#3 plots of distribution of propensity scores
-
+# plot of distribution of propensity scores 
 ggplot(df, aes(x = ps, group = inanalysislabel)) + geom_density(aes(fill = inanalysislabel), alpha = 0.2) +
@@
 summ_dist <- summary(model_dist)
+```
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # Model Summaries (Note coefficients are per 1000 women)
@@
                   "p-value" = c('< 0.001')) %>% 
-  kbl(caption = "Table 3: Change in Abortions per 1000 Women in Model without Distances", 
+  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
           digits = 3)
-print(table_3)
+table_3
+```
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 p <- summ_dist$coefficients[4,4] %>% round(3)
@@
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%
-  kbl(caption = "Table 4: Change in Abortions per 1000 Women in Model with Distances")
-print(table_4)
+  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances")
+table_4
 
@@
 propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-summary(propensitywts)
+# summary(propensitywts)
 
@@
 
-#3 plots of distribution of propensity scores
-
+# plot of distribution of propensity scores 
 ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
@@
   ylab("Density")
-
-ggplot(df, aes(x = ps, fill = hasabortiondatalabel, color = hasabortiondatalabel)) +
-  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
-  facet_grid(as.factor(hasabortiondata) ~ .) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  scale_color_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical')
-
-df %>%
-  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
-  group_by(hasabortiondata, ps.grp) %>%
-  summarize(n = n()) %>%
-  ungroup() %>%
-  mutate(n2 = ifelse(hasabortiondata == 0, yes = n, no =  -1*n)) %>%
-  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasabortiondata))) +
-  geom_bar(stat = 'identity', position = 'identity') +
-  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasabortiondata == 0, 8, -8))) +
-  xlab('Probability of Being in the Analysis') +
-  ylab('N') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
-  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
-  theme(legend.position = 'bottom', legend.direction = 'vertical',
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank())
 
@@
                   "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
-  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model without Distances, all 50 States",
-          digits = 3)
-print(table_ext_no_dist)
+  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances, all 50 States", digits = 3)
+table_ext_no_dist
+```
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
@@
                   "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
-  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
-print(table_ext_dist)
-
+  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
+table_ext_dist
 ```
@@
        ))
+```
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #calculate propensity scores
@@
 
-#3 plots of distribution of propensity scores
-
+# plot of distribution of propensity scores 
 plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
@@
   ylab("Density")
-print(plot1)
+plot1
 
@@
 model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
-summary(model)
+# summary(model)
 
@@
 propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-summary(propensitywts)
+# summary(propensitywts)
 
@@
 df$ps <- p.score
-#3 plots of distribution of propensity scores
+
+# plot of distribution of propensity scores
 
@@
 model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
-summary(model)
+# summary(model)
 
@@
 propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-summary(propensitywts)
+# summary(propensitywts)
 
@@
 df$ps <- p.score
-#3 plots of distribution of propensity scores
+# plot of distribution of propensity scores
 
@@
 model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
-summary(model)
+# summary(model)
 

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 REPLICATION:
-TODO: averages of averages vs. averages for table 2
-
+TODO: averages of averages for table 2
+TODO: reload countydem file with correct college grad pct.
 TODO: the paper says "all models included state and year fixed effects and standard errors adjusted for state-level clustering" -> does this mean we need to include states as a variable in our model?
 -- No, they did that because they were doing the longitudinal analysis, so like abortion rates in a given state for one year are not independent of the next, so they're accounting for that.
-
 TODO: the paper says "We addressed data gaps in demographic variables by using the estimate from the nearest year for which an estimate was available", so we need to include this methodology
@@
 TODO: write literature review and describe motivation for extension
-TODO: figure out what is happening with the propensity scores for the extension analyses
+TODO: rerun extension on state level
 TODO: decide whether to rescore the high vs. low restriction variable as a factor or an integer (currently coded as the latter)

### 2021-12-15 report.Rmd
Update report.Rmd

@@
 Tables 3 and 4 show the main results from our replication.
-Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a  reduction in abortions of 0.922 per 1,000 women, and that effect was reduced to 0.828 when adjusting for distances to providers.
+Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a  reduction in abortions of 0.814 per 1,000 women, and that effect was reduced to 0.649 when adjusting for distances to providers.
 These effects were roughly twice as large as the effects calculated by Brown et al., nevertheless our estimates were within their reported 95% confidence intervals (see Brown et al.

### 2021-12-15 paper_replication.Rmd
Resize plot

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=5, fig.align='center'}
 #calculate propensity scores

### 2021-12-15 report.Rmd
Update report.Rmd

@@
 
-In this paper we replicate an analysis published in JAMA Network Open by Benjamin P.
+In this paper we replicate and extend an analysis published in JAMA Network Open by Benjamin P.
 Brown, MD, MS; Luciana E.
 Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  Brown et al.
-use a binary classification of states legislative climates on abortion as either "highly restrictive" or not, and they assess the association with abortion rates using propensity weighted linear difference in differences models, based on data from 18 states.
+use a binary classification of state legislative climates on abortion as either "highly restrictive" or not, and they assess the association with abortion rates using propensity weighted linear difference in differences models, based on longitudinal data from 18 states.
 
+We first reproduce the analysis of Brown et al.
+using data from the same 18 states.
 The data set from the Brown et al.
@@
 
-We first reproduce the analysis of Brown et al.
-using data from the same 18 states.
-We then expand the analysis by using data from all 50 states and using an integer variable for the extent of abortion restriction legislation.
-Finally we extend the analysis to include associations of restrictive abortion legislation on  maternity mortality data, infant mortality data, and birth rate.
+We then expand the analysis by using data from all 50 states, and by using an integer variable for the extent of abortion restriction legislation.
+Finally we extend the analysis to include associations of restrictive abortion legislation with fetal mortality, infant mortality, birth rate and fertility rate.
 
@@
 
-We used propensity weighted linear models to assess the association of restrictive abortion legislation and abortion rates.
-We used data from the same set of 18 states as Brown et al.
-(see Table 1), and used propensity weighting to account for differences between those states and the other 32 states.
-The propensity weights were calculated using the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008.
-We fit linear models with and without including distances to abortion providers and an interaction term between those distances and restrictive legislation.
+In our replication, we used propensity weighted linear models to assess the association of restrictive abortion legislation and abortion rates, using data from the same set of 18 states as Brown et al.
+(see Table 1).
+We used propensity weighting to account for differences between those states and the other 32 statesbased on the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008.
+We fit linear models with and without distances to abortion providers and an interaction term between those distances and restrictive legislation.
 
@@
 
-We estimated distances to providers as the straight-line distance from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
+We estimated distances to providers as the straight-line distances from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
 
@@
 The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).
-This suggest that the 18 states included in the analysis are not exceptionally different from those not included regarding the demographic covariates, and accordingly we expect the propensity weights to accurately adjust for the differences.
+This suggests that the 18 states included in the analysis are not exceptionally different from those not included (in terms the demographic covariates listed above), and accordingly we expect the propensity weights to accurately adjust for the differences.
 

### 2021-12-15 paper_replication.Rmd
Table 2 p-values formatting

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# names(countydem)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+##### Table 2 ####
+
 countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
 
-countydem$inanalysislabel <- factor(countydem$inanalysis, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))
+countydem$inanalysislabel <- factor(countydem$inanalysis, levels = c(1, 0), labels = c("In Analysis", "Not in Analysis"))
 
@@
 label(countydem$College.Grad.pct) = "College Graduates"
-label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+label(countydem$percentdem2008) = "Vote Democratic"
 
@@
     g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+    p <- t.test(y ~ g)$p.value  %>% round(3) 
+    c("", sub("<", "<", format.pval(p, digits=3, eps=0.001)))
 }
 
+## Table 2 ###
 table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
        data = countydem,
+       overall = FALSE,
        extra.col = list(`P-value` = p_value_t_test),
+       extra.col.pos = 3,
        footnote = 'P-values are derived by two sample t-tests.',

### 2021-12-15 report.Rmd
Motivation for extension

@@
 Specifically we analyzed data from 2010, primarily because this was the year within the original paper's time range for which we had the most detailed information on state legislation restricting abortions.
+We then expand the analysis by using data from all 50 states, and by using an integer variable for the extent of abortion restriction legislation.
 
-We then expand the analysis by using data from all 50 states, and by using an integer variable for the extent of abortion restriction legislation.
-Finally we extend the analysis to include associations of restrictive abortion legislation with fetal mortality, infant mortality, birth rate and fertility rate.
+The original study addresses the immediate, intended consequences of the restrictive abortion legislation.
+In our paper, we extend the analysis to include associations of restrictive abortion legislation with fetal mortality, infant mortality, birth rate and fertility rate.
+Restricted access to abortion has been linked to adverse outcomes for mothers and children, including pre-term birth and low birth weight (Redd et al., 2021), infant mortality rates (Edwards, 2019), and maternal mortality (Stevenson, 2021).
+However, the effects of abortion legislation on these health outcomes are less well understood.
+The key findings of a meta-analysis (Austin and Harper, 2017) were that "Certain TRAP laws may be associated with a decrease in state-level abortion rates.
+Additional research is required to better understand the impact of these laws on women’s health outcomes."
 
@@
 The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.
+
+## Sources
+Austin N, Harper S. "Assessing the impact of TRAP laws on abortion and women’s health in the USA: a systematic review" BMJ Sex Reprod Health, 2018; 44:128–134.
+
+Edwards, E.
+"States pushing abortion bans have higher infant mortality rates".
+NBC News Website, 2019.
+Available at https://www.nbcnews.com/health/womens-health/states-pushing-abortion-bans-have-higher-infant-mortality-rates-n1008481.
+Accessed 12/10/21.
+
+Redd, Sara K.; Stidham Hall, Kelli; Aswani, Monica S.; Sen, Bisakha; Wingate, Martha; Rice, Whitney S.
+"Variation in Restrictive Abortion Policies and Adverse Birth Outcomes in the United States from 2005 to 2015" Women's Health Issues, 2021 1-11.
+
+Stevenson, A. J. "The Pregnancy-Related Mortality Impact of a Total Abortion Ban in the United States: A Research Note on Increased
+Deaths Due to Remaining Pregnant" Demography (2021) 58(6):2019–2028.

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

@@
 countydem$county <- str_to_title(countydem$county)
+countydem$county <- gsub(countydem$county, pattern = "Doña Ana", replacement = "Dona Ana")
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r}
 ####### add in natality, maternalmortality, and infantdeath data #######
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r}
 #### recode restriction level and include all counties with abortion data in analysis ####
+countyanalysis <- function(varname) {
+  countydem <- countydem %>% mutate(hasdata = case_when(is.na(countydem[[varname]]) ~ 0, 
+                                                              !(is.na(countydem[[varname]])) ~ 1))
+  countydemanalysis <- countydem %>% filter(hasdata == 1)
+  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+  usedcounties <- aggregate(countydem[[varname]] ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+  summarytable <- left_join(summarytable, usedcounties, by = "state")
+  print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Data Included in Analysis by State") %>% kable_styling())
 
-countydem <- countydem %>% mutate(hasabortiondata = case_when(is.na(AbortionRate2010) ~ 0, 
-                                                              !(is.na(AbortionRate2010)) ~ 1))
-
-countydemanalysis <- countydem %>% filter(hasabortiondata == 1)
-
-summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
-usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-summarytable <- left_join(summarytable, usedcounties, by = "state")
-summarytable
-summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
-# names(countydem)
-countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
-
-countydem$hasabortiondatalabel <- factor(countydem$hasabortiondata, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))
-
-label(countydem$Women) = "Women"
-label(countydem$White.pct) = "White"
-label(countydem$Black.pct) = "Black"
-label(countydem$Native.American.pct) = "Native American"
-label(countydem$Asian.pct) = "Asian"
-label(countydem$Hispanic.pct) = "Hispanic"
-label(countydem$Median.Income) = "Median Income"
-label(countydem$College.Grad.pct) = "College Graduates"
-label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
-
-units(countydem$Women) = "No."
-units(countydem$White.pct) = "%"
-units(countydem$Black.pct) = "%"
-units(countydem$Native.American.pct) = "%"
-units(countydem$Asian.pct) = "%"
-units(countydem$Hispanic.pct) = "%"
-units(countydem$Median.Income) = "$"
-units(countydem$College.Grad.pct) = "%"
-units(countydem$percentdem2008) = "%"
-
-# The following function is based on 
-# "Using the table1 Package to Create HTML Tables of Descriptive Statistics"
-# by Benjamin Rich.  Available at:
-# https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html#example-a-column-of-p-values
-p_value_t_test <- function(x, ...){
-    y <- unlist(x[1:2])
-    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
-}
-
-table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasabortiondatalabel, 
-       data = countydem,
-       extra.col = list(`P-value` = p_value_t_test),
-       footnote = 'P-values are derived by two sample t-tests.'
-       )
-
-#calculate propensity scores
-#include education this time
-propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-# summary(propensitywts)
-
-p.score <- predict(propensitywts, countydem, type = "response")
-df <- countydem
-df$ps <- p.score
-
-# plot of distribution of propensity scores 
-ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  ylab("Density")
-
-#run models, all 50 states, integer restrictions
-countydemanalysis <- df %>% filter(hasabortiondata == 1)
-
-ext_model_dist <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
-summ_ext_dist <- summary(ext_model_dist)
-
-ext_model_no_dist <- lm(AbortionRate2010 ~ totalrestrict, data = countydemanalysis, weights = ps)
-summ_ext_no_dist <- summary(ext_model_no_dist)
-
-# Model Summaries (Note coefficients are per 1000 women)
-table_ext_no_dist <- tibble(Characteristic = "Number of Types of Restrictive Legislation",
-                  Beta = ext_model_no_dist$coefficients[2]*1000,
-                  "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
-  kbl(caption = "Change in Abortions per 1000 Women in Model without Distances, all 50 States", digits = 3)
-table_ext_no_dist
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
-                                            "Distance to Provider", 
-                                            "Distance/Restriction Interaction"),
-                  Beta = ext_model_dist$coefficients[2:4]*1000,
-                  "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
-  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
-table_ext_dist
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
-countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
-countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
-countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
-hist(countydem$maternaldeathrate)
-
-countydem$infantdeathrate <- na_if(countydem$infantdeathrate, "Missing County")
-countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "[()]", replacement = "")
-countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "Unreliable", replacement = "")
-countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
-hist(countydem$infantdeathrate)
-
-countydem$Birth.Rate <- na_if(countydem$Birth.Rate, "Missing County")
-countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
-hist(countydem$Birth.Rate)
-
-countydem$Fertility.Rate <- na_if(countydem$Fertility.Rate, "Missing County")
-countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
-hist(countydem$Fertility.Rate)
-
-```
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-### repeat analysis for infant death rate ###
-countydem <- countydem %>% mutate(hasinfantdeathdata = case_when(is.na(countydem$infantdeathrate) ~ 0, 
-                                                              !(is.na(countydem$infantdeathrate)) ~ 1))
-
-countydeminfantdeathanalysis <- countydem %>% filter(hasinfantdeathdata == 1)
-
-summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
-usedcounties <- aggregate(infantdeathrate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-summarytable <- left_join(summarytable, usedcounties, by = "state")
-summarytable %>% mutate(percent = infantdeathrate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Data Included in Analysis by State") %>% kable_styling()
-# names(countydem)
-countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
-
-countydem$hasinfantdeathdatalabel <- factor(countydem$hasinfantdeathdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
+  countydem$hasdatalabel <- factor(countydem$hasdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
 
@@
 
-print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasinfantdeathdatalabel, 
+  print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
        data = countydem,
@@
        ))
-```
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
   #calculate propensity scores
   #include education this time
-propensitywts <- glm(hasinfantdeathdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
+  propensitywts <- glm(hasdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
 
@@
 
-# plot of distribution of propensity scores 
-plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
+  #plot of distribution of propensity scores 
+  p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
   xlab('Probability of Being in the Analysis') +
@@
   ylab("Density")
-plot1
+  print(p1)
 
   #run model
-countydeminfantdeathanalysis <- df %>% filter(hasinfantdeathdata == 1)
+  countydemanalysis <- df %>% filter(hasdata == 1)
 
-model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
-# summary(model)
+  model <- lm(countydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
+  return(summary(model))
+}
+
+countyanalysis("AbortionRate2010")
+```
+
+```{r}
+###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
+countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
+countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
+countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
+hist(countydem$maternaldeathrate)
+
+countydem$infantdeathrate <- na_if(countydem$infantdeathrate, "Missing County")
+countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "[()]", replacement = "")
+countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "Unreliable", replacement = "")
+countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
+hist(countydem$infantdeathrate)
+
+countydem$Birth.Rate <- na_if(countydem$Birth.Rate, "Missing County")
+countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
+hist(countydem$Birth.Rate)
+
+countydem$Fertility.Rate <- na_if(countydem$Fertility.Rate, "Missing County")
+countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
+hist(countydem$Fertility.Rate)
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-### repeat analysis for birth rate ###
+```{r}
 
-countydem <- countydem %>% mutate(hasbirthratedata = case_when(is.na(countydem$Birth.Rate) ~ 0, 
-                                                              !(is.na(countydem$Birth.Rate)) ~ 1))
+largecountyanalysis <- function(varname) {
+  largecountydem <- countydem %>% filter(Total >= 250000)
+  largecountydem <- largecountydem %>% mutate(hasdata = case_when(is.na(largecountydem[[varname]]) ~ 0, 
+                                                              !(is.na(largecountydem[[varname]])) ~ 1))
+  largecountydemanalysis <- largecountydem %>% filter(hasdata == 1)
+  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
+  usedcounties <- aggregate(largecountydem[[varname]] ~ state, data = largecountydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
+  summarytable <- left_join(summarytable, usedcounties, by = "state")
+  print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+Number of Counties with Data Included in Analysis by State") %>% kable_styling())
 
-countydembirthrateanalysis <- countydem %>% filter(hasbirthratedata == 1)
+  largecountydem$hasdatalabel <- factor(largecountydem$hasdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
 
-summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
-usedcounties <- aggregate(Birth.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-usedcounties
-summarytable <- left_join(summarytable, usedcounties, by = "state")
-summarytable
-summarytable %>% mutate(percent = Birth.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Data Included in Analysis by State") %>% kable_styling()
-# names(countydem)
-countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
+  label(largecountydem$Women) = "Women"
+  label(largecountydem$White.pct) = "White"
+  label(largecountydem$Black.pct) = "Black"
+  label(largecountydem$Native.American.pct) = "Native American"
+  label(largecountydem$Asian.pct) = "Asian"
+  label(largecountydem$Hispanic.pct) = "Hispanic"
+  label(largecountydem$Median.Income) = "Median Income"
+  label(largecountydem$College.Grad.pct) = "College Graduates"
+  label(largecountydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
 
-countydem$hasbirthratedatalabel <- factor(countydem$hasbirthratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
+  units(largecountydem$Women) = "No."
+  units(largecountydem$White.pct) = "%"
+  units(largecountydem$Black.pct) = "%"
+  units(largecountydem$Native.American.pct) = "%"
+  units(largecountydem$Asian.pct) = "%"
+  units(largecountydem$Hispanic.pct) = "%"
+  units(largecountydem$Median.Income) = "$"
+  units(largecountydem$College.Grad.pct) = "%"
+  units(largecountydem$percentdem2008) = "%"
 
-label(countydem$Women) = "Women"
-label(countydem$White.pct) = "White"
-label(countydem$Black.pct) = "Black"
-label(countydem$Native.American.pct) = "Native American"
-label(countydem$Asian.pct) = "Asian"
-label(countydem$Hispanic.pct) = "Hispanic"
-label(countydem$Median.Income) = "Median Income"
-label(countydem$College.Grad.pct) = "College Graduates"
-label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
+  print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
+       data = largecountydem,
+       ))
 
-units(countydem$Women) = "No."
-units(countydem$White.pct) = "%"
-units(countydem$Black.pct) = "%"
-units(countydem$Native.American.pct) = "%"
-units(countydem$Asian.pct) = "%"
-units(countydem$Hispanic.pct) = "%"
-units(countydem$Median.Income) = "$"
-units(countydem$College.Grad.pct) = "%"
-units(countydem$percentdem2008) = "%"
+  #calculate propensity scores
+  #include education this time
+  propensitywts <- glm(hasdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=largecountydem, family=binomial())
+
+  p.score <- predict(propensitywts, largecountydem, type = "response")
+  df <- largecountydem
+  df$ps <- p.score
+
+  #plot of distribution of propensity scores 
+  p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
+  xlab('Probability of Being in the Analysis') +
+  ggtitle('Propensity Score Distribution by Group') +
+  scale_fill_discrete('') +
+  theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  ylab("Density")
+  print(p1)
+
+  #run model
+  largecountydemanalysis <- df %>% filter(hasdata == 1)
+
+  model <- lm(largecountydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = largecountydemanalysis, weights = ps)
+  return(summary(model))
+}
+
+largecountyanalysis("infantdeathrate")
+
+```
+
+```{r}
+largecountyanalysis("infantdeathrate")
+```
+
+```{r}
+largecountyanalysis("Birth.Rate")
+```
+
+```{r}
+largecountyanalysis("Fertility.Rate")
+```
+
+```{r}
+#add in state level maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
+
+statelevelnatality <- read.delim("statelevelnatality.txt")
+statelevelinfantdeaths <- read.delim("statelevelinfantdeaths.txt")
+statelevelfetaldeaths <- read.delim("statelevelfetaldeaths.txt")
+statelevelmaternalmortality <- read.delim("statelevelmaternalmortality.txt")
+
+statelevelnatality <- statelevelnatality %>% select(State, Birth.Rate, Fertility.Rate)
+statelevelinfantdeaths <- statelevelinfantdeaths %>% select(State, Death.Rate)
+statelevelfetaldeaths <- statelevelfetaldeaths %>% select(Standard.Residence.States, Percent.of.Total.Deaths) %>% rename(State = Standard.Residence.States, Pct.Fetal.Deaths = Percent.of.Total.Deaths)
+statelevelmaternalmortality <- statelevelmaternalmortality %>% select(State, X..of.Total.Deaths) %>% rename(Pct.Maternal.Deaths = X..of.Total.Deaths)
+
+statedem <- countydem %>% group_by(state) %>% summarize(AvgWomen = mean(Women, na.rm = TRUE), AvgWhite.pct = mean(White.pct, na.rm = TRUE), AvgBlack.pct = mean(Black.pct, na.rm = TRUE), AvgNative.American.pct = mean(Native.American.pct, na.rm = TRUE), AvgAsian.pct = mean(Asian.pct, na.rm = TRUE), AvgHispanic.pct = mean(Hispanic.pct, na.rm = TRUE), AvgCollege.Grad.pct = mean(College.Grad.pct, na.rm = TRUE), AvgMedian.Income = mean(Median.Income, na.rm = TRUE), AvgDem2008.pct = mean(percentdem2008, na.rm = TRUE), AvgDistancetoProvider = mean(distancetoprovider, na.rm = TRUE), TotalRestrictions = mean(totalrestrict, na.rm = TRUE), AvgAbortionRate2010 = mean(AbortionRate2010, na.rm = TRUE)) %>% rename(State = state)
+
+statedem <- left_join(statedem, statelevelnatality, by = c("State"))
+statedem <- left_join(statedem, statelevelinfantdeaths , by = c("State"))
+statedem <- left_join(statedem, statelevelfetaldeaths, by = c("State"))
+statedem <- left_join(statedem, statelevelmaternalmortality, by = c("State"))
+
+for (i in 1:ncol(statedem)) {
+  statedem[[i]][is.nan(statedem[[i]])] <- NA
+}
+
+statedem$Pct.Maternal.Deaths <- na_if(statedem$Pct.Maternal.Deaths, "Suppressed")
+statedem$Pct.Maternal.Deaths <- gsub(statedem$Pct.Maternal.Deaths, pattern = "%", replacement = "")
+statedem$Pct.Maternal.Deaths <- as.numeric(statedem$Pct.Maternal.Deaths)
+
+```
+
+```{r}
+###make function for extended analysis###
+stateanalysis <- function(varname) {
+  statedem <- statedem %>% mutate(hasdata = case_when(is.na(statedem[[varname]]) ~ 0, 
+                                                              !(is.na(statedem[[varname]])) ~ 1))
+  statedemanalysis <- statedem %>% filter(hasdata == 1)
+  statedem$hasdatalabel <- factor(statedem$hasdata, levels = c(1, 0), labels = c("States With Data", "States Without Data"))
+
+  label(statedem$AvgWomen) = "Average Women"
+  label(statedem$AvgWhite.pct) = "Average White"
+  label(statedem$AvgBlack.pct) = "Average Black"
+  label(statedem$AvgNative.American.pct) = "Average Native American"
+  label(statedem$AvgAsian.pct) = "Average Asian"
+  label(statedem$AvgHispanic.pct) = "Average Hispanic"
+  label(statedem$AvgMedian.Income) = "Average Median Income"
+  label(statedem$AvgCollege.Grad.pct) = "Average College Graduates"
+  label(statedem$AvgDem2008.pct) = "Average Voting for the Democratic candidate in the most recent presidential election"
+
+  units(statedem$AvgWomen) = "No."
+  units(statedem$AvgWhite.pct) = "%"
+  units(statedem$AvgBlack.pct) = "%"
+  units(statedem$AvgNative.American.pct) = "%"
+  units(statedem$AvgAsian.pct) = "%"
+  units(statedem$AvgHispanic.pct) = "%"
+  units(statedem$AvgMedian.Income) = "$"
+  units(statedem$AvgCollege.Grad.pct) = "%"
+  units(statedem$AvgDem2008.pct) = "%"
 
@@
 
-print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasbirthratedatalabel, 
-       data = countydem,
+  print(table1(~ AvgWomen + AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct | hasdatalabel, 
+       data = statedem,
        extra.col = list(`P-value` = p_value_t_test),
@@
   #include education this time
-propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-# summary(propensitywts)
+  propensitywts <- glm(hasdata ~ AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct, data=statedem, family=binomial())
 
-p.score <- predict(propensitywts, countydem, type = "response")
-df <- countydem
+  p.score <- predict(propensitywts, statedem, type = "response")
+  df <- statedem
   df$ps <- p.score
 
-# plot of distribution of propensity scores
+  #plot of distribution of propensity scores
 
-plot1 <- ggplot(df, aes(x = ps, group = hasbirthratedatalabel)) + geom_density(aes(fill = hasbirthratedatalabel), alpha = 0.2) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  ylab("Density")
-plot1
-
-#run model
-countydembirthrateanalysis <- df %>% filter(hasbirthratedata == 1)
-
-model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
-# summary(model)
-
-test <- countydembirthrateanalysis %>% filter(ps > 0.9999999)
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-### repeat analysis for fertility rate ###
-
-countydem <- countydem %>% mutate(hasfertilityratedata = case_when(is.na(countydem$Fertility.Rate) ~ 0, 
-                                                              !(is.na(countydem$Fertility.Rate)) ~ 1))
-
-countydemfertilityrateanalysis <- countydem %>% filter(hasfertilityratedata == 1)
-
-summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
-usedcounties <- aggregate(Fertility.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-usedcounties
-summarytable <- left_join(summarytable, usedcounties, by = "state")
-summarytable
-summarytable %>% mutate(percent = Fertility.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Data Included in Analysis by State") %>% kable_styling()
-# names(countydem)
-countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)
-
-countydem$hasfertilityratedatalabel <- factor(countydem$hasfertilityratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
-
-label(countydem$Women) = "Women"
-label(countydem$White.pct) = "White"
-label(countydem$Black.pct) = "Black"
-label(countydem$Native.American.pct) = "Native American"
-label(countydem$Asian.pct) = "Asian"
-label(countydem$Hispanic.pct) = "Hispanic"
-label(countydem$Median.Income) = "Median Income"
-label(countydem$College.Grad.pct) = "College Graduates"
-label(countydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
-
-units(countydem$Women) = "No."
-units(countydem$White.pct) = "%"
-units(countydem$Black.pct) = "%"
-units(countydem$Native.American.pct) = "%"
-units(countydem$Asian.pct) = "%"
-units(countydem$Hispanic.pct) = "%"
-units(countydem$Median.Income) = "$"
-units(countydem$College.Grad.pct) = "%"
-units(countydem$percentdem2008) = "%"
-
-p_value_t_test <- function(x, ...){
-    y <- unlist(x[1:2])
-    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
-}
-
-print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasfertilityratedatalabel, 
-       data = countydem,
-       extra.col = list(`P-value` = p_value_t_test),
-       footnote = 'P-values are derived by two sample t-tests.'
-       ))
-
-#calculate propensity scores
-#include education this time
-propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
-# summary(propensitywts)
-
-p.score <- predict(propensitywts, countydem, type = "response")
-df <- countydem
-df$ps <- p.score
-# plot of distribution of propensity scores
-
-plot1 <- ggplot(df, aes(x = ps, group = hasfertilityratedatalabel)) + geom_density(aes(fill = hasfertilityratedatalabel), alpha = 0.2) +
+  plot1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
     xlab('Probability of Being in the Analysis') +
@@
   #run model
-countydemfertilityrateanalysis <- df %>% filter(hasfertilityratedata == 1)
+  statedemanalysis <- df %>% filter(hasdata == 1)
 
-model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
-# summary(model)
+  model <- lm(statedemanalysis[[varname]] ~ TotalRestrictions + AvgDistancetoProvider, data = statedemanalysis, weights = ps)
+  summary(model)
+}
 
+```
+
+```{r}
+stateanalysis("AvgAbortionRate2010")
+```
+
+```{r}
+stateanalysis("Birth.Rate")
+```
+
+```{r}
+stateanalysis("Fertility.Rate")
+```
+
+```{r}
+stateanalysis("Death.Rate")
+```
+
+```{r}
+stateanalysis("Pct.Fetal.Deaths")
+```
+
+```{r}
+stateanalysis("Pct.Maternal.Deaths")
 ```

### 2021-12-15 report.Rmd
Update report.Rmd

@@
 Restricted access to abortion has been linked to adverse outcomes for mothers and children, including pre-term birth and low birth weight (Redd et al., 2021), infant mortality rates (Edwards, 2019), and maternal mortality (Stevenson, 2021).
-However, the effects of abortion legislation on these health outcomes are less well understood.
-The key findings of a meta-analysis (Austin and Harper, 2017) were that "Certain TRAP laws may be associated with a decrease in state-level abortion rates.
-Additional research is required to better understand the impact of these laws on women’s health outcomes."
+However, the effects of abortion legislation on these health outcomes are less well understood than the direct effect on number of abortions.
+The key findings of a meta-analysis on the effects of targeted restrictions on abortion providers (TRAP) were that "Certain TRAP laws may be associated with a decrease in state-level abortion rates" and "Additional research is required to better understand the impact of these laws on women’s health outcomes." (Austin and Harper, 2017)  We attempt to address this needd in the existing literature by adapting the methodology of Brown et al.
+to quantify the health impacts of restrictive abortion legislation.
+We also measure the association of these laws on birth rates, as restrictions on abortion have in some cases been promoted as an antedote to declining birth rates, recently in eastern Europe (Nikolova, 2020).
 
@@
 
-For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, targeted restrictions on abortion providers (TRAP) laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors.
+For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, TRAP laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors.
 Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3/4 types of laws.
@@
 Available at https://www.nbcnews.com/health/womens-health/states-pushing-abortion-bans-have-higher-infant-mortality-rates-n1008481.
-Accessed 12/10/21.
+Accessed 12/14/21.
+
+Nikolova, M.
+"When will conservatives learn that banning abortion does little to address demographic decline?" Emerging Europe website, 2020.
+Available at https://emerging-europe.com/after-hours/when-will-conservatives-learn-that-banning-abortion-does-little-to-address-demographic-decline/.
+Accessed 12/14/21.
 

### 2021-12-16 paper_replication.Rmd
All the tables

@@
 
-to_adjust_kable_table_numbers <- kable(table_1)
 ```
@@
 
+###### End of Replication ######
 ```
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#### Begin Extension #####
 ####### add in natality, maternalmortality, and infantdeath data #######
@@
 infantdeath$County <- gsub(infantdeath$County, pattern = " City", replacement = "")
-infantdeath$County[!(infantdeath$County %in% countydem$county)]
+# infantdeath$County[!(infantdeath$County %in% countydem$county)]
 infantdeath <- infantdeath %>% rename(state = State, county = County, infantdeathrate = Death.Rate)
@@
 maternalmortality$County[!(maternalmortality$County %in% countydem$county)] <- gsub(maternalmortality$County[!(maternalmortality$County %in% countydem$county)], pattern = " And", replacement = "")
-maternalmortality$County[!(maternalmortality$County %in% countydem$county)]
+# maternalmortality$County[!(maternalmortality$County %in% countydem$county)]
 
@@
 natality$County <- gsub(natality$County, pattern = "La Porte", replacement = "Laporte")
-natality$County[!(natality$County %in% countydem$county)]
+# natality$County[!(natality$County %in% countydem$county)]
 natality <- natality %>% rename(state = State, county = County)
@@
 
-sum(is.na(countydem$infantdeathrate)/nrow(countydem))
+# sum(is.na(countydem$infantdeathrate)/nrow(countydem))
 countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
@@
 
-sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
-sum(is.na(countydem$Birth.Rate)/nrow(countydem))
+# sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
+# sum(is.na(countydem$Birth.Rate)/nrow(countydem))
 
 ```
-
-```{r}
+## Expansion to Data from All States
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #### recode restriction level and include all counties with abortion data in analysis ####
@@
   summarytable <- left_join(summarytable, usedcounties, by = "state")
-  print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Data Included in Analysis by State") %>% kable_styling())
+  summarytable <- summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% 
+    kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), 
+        caption = "Number of Counties with Data Included in Analysis by State") %>% 
+    kable_styling()
 
@@
 
-  p_value_t_test <- function(x, ...){
-    y <- unlist(x[1:2])
-    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
-  }
-
-  print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
-       data = countydem,
-       extra.col = list(`P-value` = p_value_t_test),
-       footnote = 'P-values are derived by two sample t-tests.'
-       ))
+  # p_value_t_test <- function(x, ...){
+  #   y <- unlist(x[1:2])
+  #   g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
+  #   p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
+  #   p_value <- ifelse(p_value == '0', '< 0.001', p_value)
+  # }
+  # 
+  # in_analysis_comp_table <- table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
+  #      data = countydem,
+  #      extra.col = list(`P-value` = p_value_t_test),
+  #      footnote = 'P-values are derived by two sample t-tests.'
+  #      )
 
@@
   #plot of distribution of propensity scores 
-  p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  ylab("Density")
-  print(p1)
+  # p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
+  # xlab('Probability of Being in the Analysis') +
+  # ggtitle('Propensity Score Distribution by Group') +
+  # scale_fill_discrete('') +
+  # theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  # ylab("Density")
+  # print(p1)
 
@@
   countydemanalysis <- df %>% filter(hasdata == 1)
+  model_no_dist <- lm(countydemanalysis[[varname]] ~ totalrestrict, data = countydemanalysis, weights = ps)
 
-  model <- lm(countydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
-  return(summary(model))
+  model_dist <- lm(countydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
+  return(list(summary(model_no_dist), summary(model_dist), summarytable))
 }
 
-countyanalysis("AbortionRate2010")
+returns <- countyanalysis("AbortionRate2010")
+all_state_model_no_dist <- returns[[1]]
+all_state_model_dist <- returns[[2]]
+
+# Model Summaries (Note coefficients are per 1000 women)
+table_5 <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = all_state_model_no_dist$coefficients[2]*1000,
+                  "p-value" = c('< 0.001')) %>% 
+  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
+          digits = 3)
+table_5
+
+p <- all_state_model_dist$coefficients[4,4] %>% round(3)
+table_6 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(all_state_model_dist$coefficients[2]*1000, 
+                           all_state_model_dist$coefficients[3]*1000,
+                           all_state_model_dist$coefficients[4]*1000),
+                  "p-value" = c('< 0.001', '< 0.001', p)) %>%
+  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances")
+table_6
+
+all_state_table <- returns[[3]]
+# all_state_table # move to appendix?
 ```
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
@@
 countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
-hist(countydem$maternaldeathrate)
+# hist(countydem$maternaldeathrate)
 
@@
 countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
-hist(countydem$infantdeathrate)
+# hist(countydem$infantdeathrate)
 
@@
 countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
-hist(countydem$Birth.Rate)
+# hist(countydem$Birth.Rate)
 
@@
 countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
-hist(countydem$Fertility.Rate)
+# hist(countydem$Fertility.Rate)
 
@@
 
-```{r}
+## Extesion to Infant Death Rate, Birth Rate, and Fertility Rate
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#### Large County Function
 
@@
   summarytable <- left_join(summarytable, usedcounties, by = "state")
-  print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
+  # print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
 Number of Counties with Data Included in Analysis by State") %>% kable_styling())
@@
 
-  print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
-       data = largecountydem,
-       ))
+  # print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
+  #      data = largecountydem,
+  #     ))
 
@@
   ylab("Density")
-  print(p1)
+  # print(p1)
 
@@
 
-  model <- lm(largecountydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = largecountydemanalysis, weights = ps)
-  return(summary(model))
+  model_no_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict, data = largecountydemanalysis, weights = ps)
+
+  model_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = largecountydemanalysis, weights = ps)
+
+  return(list(summary(model_no_dist), summary(model_dist)))
 }
 
-largecountyanalysis("infantdeathrate")
+returns <- largecountyanalysis("infantdeathrate")
+large_county_no_dist <- returns[[1]]
+large_county_dist <- returns[[2]]
+
+# Model Summaries
+table_7 <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = large_county_no_dist$coefficients[2],
+                  "p-value" = c('< 0.001')) %>% 
+  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
+          digits = 4)
+table_7
+
+p_2 <- large_county_dist$coefficients[3,4] %>% round(3)
+p_3 <- large_county_dist$coefficients[4,4] %>% round(3)
+
+table_8 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(large_county_dist$coefficients[2], 
+                           large_county_dist$coefficients[3],
+                           large_county_dist$coefficients[4]),
+                  "p-value" = c('< 0.001', p_2, p_3)) %>%
+  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
+table_8
 
@@
 
-```{r}
-largecountyanalysis("infantdeathrate")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# largecountyanalysis("Birth.Rate")
 ```
 
-```{r}
-largecountyanalysis("Birth.Rate")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# largecountyanalysis("Fertility.Rate")
 ```
 
-```{r}
-largecountyanalysis("Fertility.Rate")
-```
-
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #add in state level maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 ###make function for extended analysis###
@@
   label(statedem$AvgCollege.Grad.pct) = "Average College Graduates"
-  label(statedem$AvgDem2008.pct) = "Average Voting for the Democratic candidate in the most recent presidential election"
+  label(statedem$AvgDem2008.pct) = "Democratic Vote"
 
@@
 
-  p_value_t_test <- function(x, ...){
-    y <- unlist(x[1:2])
-    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
-  }
-
-  print(table1(~ AvgWomen + AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct | hasdatalabel, 
-       data = statedem,
-       extra.col = list(`P-value` = p_value_t_test),
-       footnote = 'P-values are derived by two sample t-tests.'
-       ))
+  # print(table1(~ AvgWomen + AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct | hasdatalabel, 
+  #      data = statedem,
+  #      extra.col = list(`P-value` = p_value_t_test),
+  #      footnote = 'P-values are derived by two sample t-tests.'
+  #      ))
 
@@
 
-  plot1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
-    xlab('Probability of Being in the Analysis') +
-    ggtitle('Propensity Score Distribution by Group') +
-    scale_fill_discrete('') +
-    theme(legend.position = 'bottom', legend.direction = 'vertical') +
-    ylab("Density")
-  print(plot1)
+  # plot1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
+  #   xlab('Probability of Being in the Analysis') +
+  #   ggtitle('Propensity Score Distribution by Group') +
+  #   scale_fill_discrete('') +
+  #   theme(legend.position = 'bottom', legend.direction = 'vertical') +
+  #   ylab("Density")
+  # print(plot1)
 
@@
 
-  model <- lm(statedemanalysis[[varname]] ~ TotalRestrictions + AvgDistancetoProvider, data = statedemanalysis, weights = ps)
-  summary(model)
+  model_no_dist <- lm(statedemanalysis[[varname]] ~ TotalRestrictions, weights = ps, data = statedemanalysis)
+
+  model_dist <- lm(statedemanalysis[[varname]] ~ TotalRestrictions + AvgDistancetoProvider + TotalRestrictions*AvgDistancetoProvider, data = statedemanalysis, weights = ps)
+
+  return(list(summary(model_no_dist), summary(model_dist)))
 }
@@
 
-```{r}
-stateanalysis("AvgAbortionRate2010")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# returns <- stateanalysis("AvgAbortionRate2010")
 ```
 
-```{r}
-stateanalysis("Birth.Rate")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+### State Birth Rate
+returns <- stateanalysis("Birth.Rate")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_9 <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = c(p)) %>% 
+  kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
+          digits = 4)
+table_9
+
+p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_10 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c(p_1, p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances")
+table_10
 ```
 
-```{r}
-stateanalysis("Fertility.Rate")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+### State Fertility Rate
+returns <- stateanalysis("Fertility.Rate")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = c(p)) %>% 
+  kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
+          digits = 4)
+table_no_dist
+
+p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c(p_1, p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances")
+table_dist
 ```
 
-```{r}
-stateanalysis("Death.Rate")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+### State  Infant Death Rate
+returns <- stateanalysis("Death.Rate")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+# p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = '< 0.001') %>% 
+  kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
+          digits = 4)
+table_no_dist
+
+# p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c('< 0.001', p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
+table_dist
 ```
 
-```{r}
-stateanalysis("Pct.Fetal.Deaths")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+
+### State Fetal Death Percent
+returns <- stateanalysis("Pct.Fetal.Deaths")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = p) %>% 
+  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
+          digits = 4)
+table_no_dist
+
+p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c(p_1, p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances")
+table_dist
 ```
 
-```{r}
-stateanalysis("Pct.Maternal.Deaths")
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+### State Maternal Deaths
+returns <- stateanalysis("Pct.Maternal.Deaths")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = p) %>% 
+  kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
+          digits = 4)
+table_no_dist
+
+p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c(p_1, p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances")
+table_dist
 ```

### 2021-12-16 paper_replication.Rmd
kable hold_position

@@
 #load in relevant data
-providers <- read.csv("provider_list_geo.csv")
-countycentroids <- read.csv("county_pop_centroids.csv")
-countydem <- read.csv("county_demographics.csv")
-abortions <- read.csv("cleanedabortiondata.csv")
-restrictions <- read.csv("state_restrictions2010.csv")
-countypres <- read.csv("countypres.csv")
+providers <- read.csv("data/provider_list_geo.csv")
+countycentroids <- read.csv("data/county_pop_centroids.csv")
+countydem <- read.csv("data/county_demographics.csv")
+abortions <- read.csv("data/cleanedabortiondata.csv")
+restrictions <- read.csv("data/state_restrictions2010.csv")
+countypres <- read.csv("data/countypres.csv")
 ```
@@
 ########### clean county dem data ###########
-countydem <- read.csv("county_demographics.csv")
+countydem <- read.csv("data/county_demographics.csv")
 
@@
 table_1 <- left_join(table_1, usedcounties, by = "state")
-table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling()
+table_1 %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Number of Counties with Abortion Data Included in Analyses by State", digits = 3) %>% kable_styling(latex_options = 'hold_position')
 
@@
   kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
-          digits = 3)
+          digits = 3) %>%
+  kable_styling(latex_options = 'hold_position')
 table_3
@@
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%
-  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances")
+  kbl(caption = "Change in Abortions per 1000 Women in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_4
@@
 ####### add in natality, maternalmortality, and infantdeath data #######
-natality <- read.delim("natalitydata.txt")
-maternalmortality <- read.delim("maternalmortality.txt")
-infantdeath <- read.delim("infantdeathrecords.txt")
+natality <- read.delim("data/natalitydata.txt")
+maternalmortality <- read.delim("data/maternalmortality.txt")
+infantdeath <- read.delim("data/infantdeathrecords.txt")
 
@@
         caption = "Number of Counties with Data Included in Analysis by State") %>% 
-    kable_styling()
+    kable_styling(latex_options = 'hold_position')
 
@@
   kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
-          digits = 3)
+          digits = 3) %>%
+  kable_styling(latex_options = 'hold_position')
 table_5
@@
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%
-  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances")
+  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_6
@@
 
+## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
-
-## Extesion to Infant Death Rate, Birth Rate, and Fertility Rate
 
@@
   kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_7
@@
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
-  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
+  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_8
@@
 
-statelevelnatality <- read.delim("statelevelnatality.txt")
-statelevelinfantdeaths <- read.delim("statelevelinfantdeaths.txt")
-statelevelfetaldeaths <- read.delim("statelevelfetaldeaths.txt")
-statelevelmaternalmortality <- read.delim("statelevelmaternalmortality.txt")
+statelevelnatality <- read.delim("data/statelevelnatality.txt")
+statelevelinfantdeaths <- read.delim("data/statelevelinfantdeaths.txt")
+statelevelfetaldeaths <- read.delim("data/statelevelfetaldeaths.txt")
+statelevelmaternalmortality <- read.delim("data/statelevelmaternalmortality.txt")
 
@@
   kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_9
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances")
+  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_10
@@
   kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_no_dist
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances")
+  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_dist
@@
   kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_no_dist
@@
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances")
+  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_dist
@@
   kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_no_dist
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances")
+  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_dist
@@
   kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
-          digits = 4)
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
 table_no_dist
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances")
+  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
 table_dist

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
-REPLICATION:
-TODO: averages of averages for table 2
-TODO: reload countydem file with correct college grad pct.
-TODO: the paper says "all models included state and year fixed effects and standard errors adjusted for state-level clustering" -> does this mean we need to include states as a variable in our model?
--- No, they did that because they were doing the longitudinal analysis, so like abortion rates in a given state for one year are not independent of the next, so they're accounting for that.
-TODO: the paper says "We addressed data gaps in demographic variables by using the estimate from the nearest year for which an estimate was available", so we need to include this methodology
-TODO: bias analysis as done in the paper, comparing Guttmacher Census State Data to our State Data
-TODO: run the models without the distance variable as well as with the distance variable as the outcome, as done in the paper
-TODO: write up everything: we need to follow the format of the original paper <- Toby
-TODO: find abortion data used by Brown et al. to compare in discussion <- Toby
-
-EXTENSION:
-TODO: write literature review and describe motivation for extension
-TODO: rerun extension on state level
-TODO: decide whether to rescore the high vs. low restriction variable as a factor or an integer (currently coded as the latter)
-TODO: add in rural vs urban variable
-TODO: look at Hawkins et al paper linked in the original paper to see what mortality data they used. consider using mortality data by state.
 

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 ---
+## Abstract
+### Importance
+In the United States, many state legislatures have enacted a range of laws limiting access to abortions.  These laws limit access to healthcare.
+
+### Objectives
+To replicate the analysis of the association of abortion restrictions on abortion rates conducted by Brown et al.
+(2020), to expand the analysis to a wider data set, and to extend the analysis to associations with maternal, fetal and infant mortality rates, as well as birth and fertility rates.
+
+### Design and Particpants
+Propensity score adjusted linear regression of county level data on abortion rates, and state level data on mortality and natality rates, using data from 2010.
+
+### Exposures
+In the replication, the exposure is a highly restrictive abortion legislation as defined by 3 out 4 common types of abortion restrictions.
+For the extension, the exposure is the number of types of abortion legislation.
+
+### Main Outcomes and Measures
+In the replication and extension, county level abortion rates defined as abortions per 1,000 women per year.
+In the extension, state level fetal mortality rates ****, infant mortality rates ****, maternal mortality rates ****, birth rates defined as number of live births per 1,000 people, and fertility rates defined as number of live births per 1,000 women per year.
+
+### Results
+
+### Conclusions and Relevance
 

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 Tables 3 and 4 show the main results from our replication.
-Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a  reduction in abortions of 0.814 per 1,000 women, and that effect was reduced to 0.649 when adjusting for distances to providers.
-These effects were roughly twice as large as the effects calculated by Brown et al., nevertheless our estimates were within their reported 95% confidence intervals (see Brown et al.
+Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a fewer abortions at a rate of -0.814 per 1,000 women without adjusting for distance to providers, and -0.649 when adjusting for distances to providers.
+These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, −0.92 to −0.04) and -0.44 (95% CI, −0.85 to −0.03), but our estimates were well within their reported 95% confidence intervals (see Brown et al.
 Table 3).
+This corroborates the main result of Brown et al.
+despite the use of completely different data sources.
 

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 ### Results
+Our replication confirms the primary result of Brown et al.
+We found that highly restrictive legislation is associated with -0.814 per 1,000 women without adjusting for distance to providers, and -0.649 when adjusting for distances to providers.
+These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, −0.92 to −0.04) and -0.44 (95% CI, −0.85 to −0.03), but our estimates were well within their reported 95% confidence intervals.
 
@@
 
+### Replication using data from 18 states
 The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).
@@
 despite the use of completely different data sources.
+
+We did observe a decrease in abortion rates based on estimated distances to abortion providers.
+This decrease was statistically significant, but not practically significant at -3.4 x 10^-6 abortions per 1,000 women (see Table 4), compared to -0.02 in the analysis by Brown et al.
+
+### Expansion to data from all states
+When we repeated the analysis using all data available in all 50 states, we measured smaller but still significant reduction in the abortion rate, with associations of -0.216 and -0.361 without and with adjustment for distance to providers.
+These again fall well within the 95% confidence interval of the estimates of Brown et al., further confirming their primary result, and suggesting that the propensity score weighting from the 18 states in their analysis to the entire country may be leading to an over-estimate of the effects.
+
+### Extension to Mortality and Natality Rates
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
   long <- countycentroids$LONGITUDE[i]
-  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))
+  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))*3,958.8
   countycentroids$distancetoprovider[i] <- min(dists)

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

--- 2021-12-16 paper_replication.Rmd
@@
   long <- countycentroids$LONGITUDE[i]
-  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))
+  dists <- distHaversine(cbind(long, lat), cbind(providers$Lon, providers$Lat))*0.000621371
   countycentroids$distancetoprovider[i] <- min(dists)

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 ---
-## Abstract
 ### Importance
-In the United States, many state legislatures have enacted a range of laws limiting access to abortions.  These laws limit access to healthcare.
+In the United States, many state legislatures have enacted a range of laws limiting access to abortions.
+These laws limit access to crucial healthcare and have impacts on mortality, fertility, and birth rates across the country.
 
@@
 To replicate the analysis of the association of abortion restrictions on abortion rates conducted by Brown et al.
-(2020), to expand the analysis to a wider data set, and to extend the analysis to associations with maternal, fetal and infant mortality rates, as well as birth and fertility rates.
+(2020), though not from a longitudinal perspective, to expand the analysis to a wider data set, and to extend the analysis to associations with maternal, fetal and infant mortality rates, as well as birth and fertility rates.
 
-### Design and Particpants
-Propensity score adjusted linear regression of county level data on abortion rates, and state level data on mortality and natality rates, using data from 2010.
+### Design, Setting, and Particpants
+This study examined all publicly available county-level abortion rates in addition to state-level maternal, fetal and infant mortality rates as well as birth and fertility rates from the year 2010.
+The association between these rates, abortion legislation, and distance to nearest provider were evaluated using propensity score adjusted linear regression.
 
 ### Exposures
-In the replication, the exposure is a highly restrictive abortion legislation as defined by 3 out 4 common types of abortion restrictions.
-For the extension, the exposure is the number of types of abortion legislation.
+In the replication, the exposures are a highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and distance to facilities performing abortions (in miles).
+For the extension, the exposures are the number of types of abortion legislation and the distance to facilities performing abortion (in miles).
 
 ### Main Outcomes and Measures
-In the replication and extension, county level abortion rates defined as abortions per 1,000 women per year.
-In the extension, state level fetal mortality rates ****, infant mortality rates ****, maternal mortality rates ****, birth rates defined as number of live births per 1,000 people, and fertility rates defined as number of live births per 1,000 women per year.
+In the replication and extension, the main outcomes and measures were county level abortion rates defined as abortions per 1,000 women per year.
+In the extension, the main outcomes and measures were state level fetal mortality rates defined as fetal deaths per 1,000 total deaths, infant mortality rates defined as infant deaths within the first year of life per 1,000 births, maternal mortality rates defined as deaths due to maternal diseases per 1,000 total deaths, birth rates defined as number of live births per 1,000 people, and fertility rates defined as number of live births per 1,000 women per year.
 
@@
 Our replication confirms the primary result of Brown et al.
-We found that highly restrictive legislation is associated with -0.814 per 1,000 women without adjusting for distance to providers, and -0.649 when adjusting for distances to providers.
+(2020).
+We found that highly restrictive legislation is associated with 0.814 fewer abortions per 1,000 women without adjusting for distance to providers, and 0.649 fewer abortions per 1,000 women when adjusting for distances to providers.
 These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, −0.92 to −0.04) and -0.44 (95% CI, −0.85 to −0.03), but our estimates were well within their reported 95% confidence intervals.
@@
 Specifically we analyzed data from 2010, primarily because this was the year within the original paper's time range for which we had the most detailed information on state legislation restricting abortions.
-We then expand the analysis by using data from all 50 states, and by using an integer variable for the extent of abortion restriction legislation.
+Limitations of Brown et al.'s research included a lack of data on abortion rates in multiple states and a seemingly arbitrary classificaton of low vs.
+high legislative abortion restrictions.
+In order to improve the reliability and generalizability of their results, we chose to expand their analysis by using data from all 50 states, by using a more flexible integer variable for the extent of abortion restriction legislation in order, and by incorporating percentage of college graduates in the propensity score calculation.
 
@@
 We also measure the association of these laws on birth rates, as restrictions on abortion have in some cases been promoted as an antedote to declining birth rates, recently in eastern Europe (Nikolova, 2020).
+Findings regarding whether restrictions on abortion actually increase birth rates, even if they decrease abortion rates, would therefore have important policy implications.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 # Model Summaries (Note coefficients are per 1000 women)
-table_5 <- tibble(Characteristic = "Highly Restrictive Legislation",
+table_5 <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = all_state_model_no_dist$coefficients[2]*1000,
@@
 p <- all_state_model_dist$coefficients[4,4] %>% round(3)
-table_6 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+table_6 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                      "Distance/Restriction Interaction"),
@@
 # Model Summaries
-table_7 <- tibble(Characteristic = "Highly Restrictive Legislation",
+table_7 <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = large_county_no_dist$coefficients[2],
@@
 
-table_8 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+table_8 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                      "Distance/Restriction Interaction"),
@@
 # Model Summaries
-table_9 <- tibble(Characteristic = "Highly Restrictive Legislation",
+table_9 <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = model_no_dist$coefficients[2],
@@
 
-table_10 <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+table_10 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                      "Distance/Restriction Interaction"),
@@
 # Model Summaries
-table_no_dist <- tibble(Characteristic = "Highly Restrictive Legislation",
+table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = model_no_dist$coefficients[2],
@@
 
-table_dist <- tibble(Characteristic = c("Highly Restrictive Legislation", "Distance to Provider",
+table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
                                      "Distance/Restriction Interaction"),

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#### Large County Function
-
-largecountyanalysis <- function(varname) {
-  largecountydem <- countydem %>% filter(Total >= 250000)
-  largecountydem <- largecountydem %>% mutate(hasdata = case_when(is.na(largecountydem[[varname]]) ~ 0, 
-                                                              !(is.na(largecountydem[[varname]])) ~ 1))
-  largecountydemanalysis <- largecountydem %>% filter(hasdata == 1)
-  summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
-  usedcounties <- aggregate(largecountydem[[varname]] ~ state, data = largecountydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
-  summarytable <- left_join(summarytable, usedcounties, by = "state")
-  # print(summarytable %>% mutate(percent = summarytable[, 3]/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1.
-Number of Counties with Data Included in Analysis by State") %>% kable_styling())
-
-  largecountydem$hasdatalabel <- factor(largecountydem$hasdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))
-
-  label(largecountydem$Women) = "Women"
-  label(largecountydem$White.pct) = "White"
-  label(largecountydem$Black.pct) = "Black"
-  label(largecountydem$Native.American.pct) = "Native American"
-  label(largecountydem$Asian.pct) = "Asian"
-  label(largecountydem$Hispanic.pct) = "Hispanic"
-  label(largecountydem$Median.Income) = "Median Income"
-  label(largecountydem$College.Grad.pct) = "College Graduates"
-  label(largecountydem$percentdem2008) = "Voting for the Democratic candidate in the most recent presidential election"
-
-  units(largecountydem$Women) = "No."
-  units(largecountydem$White.pct) = "%"
-  units(largecountydem$Black.pct) = "%"
-  units(largecountydem$Native.American.pct) = "%"
-  units(largecountydem$Asian.pct) = "%"
-  units(largecountydem$Hispanic.pct) = "%"
-  units(largecountydem$Median.Income) = "$"
-  units(largecountydem$College.Grad.pct) = "%"
-  units(largecountydem$percentdem2008) = "%"
-
-  # print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
-  #      data = largecountydem,
-  #     ))
-
-  #calculate propensity scores
-  #include education this time
-  propensitywts <- glm(hasdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=largecountydem, family=binomial())
-
-  p.score <- predict(propensitywts, largecountydem, type = "response")
-  df <- largecountydem
-  df$ps <- p.score
-
-  #plot of distribution of propensity scores 
-  p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
-  xlab('Probability of Being in the Analysis') +
-  ggtitle('Propensity Score Distribution by Group') +
-  scale_fill_discrete('') +
-  theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  ylab("Density")
-  # print(p1)
-
-  #run model
-  largecountydemanalysis <- df %>% filter(hasdata == 1)
-
-  model_no_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict, data = largecountydemanalysis, weights = ps)
-
-  model_dist <- lm(largecountydemanalysis[[varname]] ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = largecountydemanalysis, weights = ps)
-
-  return(list(summary(model_no_dist), summary(model_dist)))
-}
-
-returns <- largecountyanalysis("infantdeathrate")
-large_county_no_dist <- returns[[1]]
-large_county_dist <- returns[[2]]
-
-# Model Summaries
-table_7 <- tibble(Characteristic = "Count of Types of Restrictions",
-                  Beta = large_county_no_dist$coefficients[2],
-                  "p-value" = c('< 0.001')) %>% 
-  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
-          digits = 4) %>%
-  kable_styling(latex_options = 'hold_position')
-table_7
-
-p_2 <- large_county_dist$coefficients[3,4] %>% round(3)
-p_3 <- large_county_dist$coefficients[4,4] %>% round(3)
-
-table_8 <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
-                                     "Distance/Restriction Interaction"),
-                  Beta = c(large_county_dist$coefficients[2], 
-                           large_county_dist$coefficients[3],
-                           large_county_dist$coefficients[4]),
-                  "p-value" = c('< 0.001', p_2, p_3)) %>%
-  kbl(caption = "Large County Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
-  kable_styling(latex_options = 'hold_position')
-table_8
-
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# largecountyanalysis("Birth.Rate")
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# largecountyanalysis("Fertility.Rate")
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #add in state level maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# returns <- stateanalysis("AvgAbortionRate2010")
+returns <- stateanalysis("AvgAbortionRate2010")
+model_no_dist <- returns[[1]]
+model_dist <- returns[[2]]
+
+p <- model_no_dist$coefficients[2,4] %>% round(3)
+# Model Summaries
+table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
+                  Beta = model_no_dist$coefficients[2],
+                  "p-value" = c(p)) %>% 
+  kbl(caption = "State Data: Abortion Rate per 1000 Women in Model without Distances", 
+          digits = 4) %>%
+  kable_styling(latex_options = 'hold_position')
+table_no_dist
+
+p_1 <- model_dist$coefficients[2,4] %>% round(3)
+p_2 <- model_dist$coefficients[3,4] %>% round(3)
+p_3 <- model_dist$coefficients[4,4] %>% round(3)
+
+table_dist <- tibble(Characteristic = c("Count of Types of Restrictions", "Distance to Provider",
+                                     "Distance/Restriction Interaction"),
+                  Beta = c(model_dist$coefficients[2], 
+                           model_dist$coefficients[3],
+                           model_dist$coefficients[4]),
+                  "p-value" = c(p_1, p_2, p_3)) %>%
+  kbl(caption = "State Data: Change in Abortion Rate 1000 Women in Model with Distances") %>%
+  kable_styling(latex_options = 'hold_position')
+table_dist
 ```
@@
                   "p-value" = c(p)) %>% 
-  kbl(caption = "State Data: Change in Birth Rate Rate per 1000 People in Model without Distances", 
+  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p)) %>% 
-  kbl(caption = "State Data: Change in Fertility Rate Rate per 1000 Women in Model without Distances", 
+  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = '< 0.001') %>% 
-  kbl(caption = "State Data: Change in Infant Death Rate Rate per 1000 **** in Model without Distances", 
+  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
           digits = 4) %>%

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
                   "p-value" = c(p)) %>% 
-  kbl(caption = "State Data: Abortion Rate per 1000 Women in Model without Distances", 
+  kbl(caption = "State Data: Abortion Rate per 1,000 Women, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Abortion Rate 1000 Women in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Abortion Rate per 1,000 Women in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   "p-value" = c(p)) %>% 
-  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model without Distances", 
+  kbl(caption = "State Data: Change in Birth Rate per 1,000 People, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Birth Rate per 1000 People in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Birth Rate per 1,000 People, in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   "p-value" = c(p)) %>% 
-  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model without Distances", 
+  kbl(caption = "State Data: Change in Fertility Rate per 1,000 Women, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Fertility Rate per 1000 Women in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Fertility Rate per 1,000 Women, in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   "p-value" = '< 0.001') %>% 
-  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model without Distances", 
+  kbl(caption = "State Data: Change in Infant Deaths per 1,000 Total Deaths, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Infant Death Rate per 1000 **** in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Infant Deaths per 1,000 Total Deaths, in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   "p-value" = p) %>% 
-  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model without Distances", 
+  kbl(caption = "State Data: Change in Fetal Deaths per 1,000 Total Deaths, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Percent Fetal Deaths in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Fetal Deaths per 1,000 Total Deaths, in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   "p-value" = p) %>% 
-  kbl(caption = "State Data: Change in Maternal Deaths per ***** in Model without Distances", 
+  kbl(caption = "State Data: Change in Maternal Deaths per 1,000 Total Deaths, in Model without Distances", 
           digits = 4) %>%
@@
                   "p-value" = c(p_1, p_2, p_3)) %>%
-  kbl(caption = "State Data: Change in Percent Maternal Deaths per **** in Model with Distances") %>%
+  kbl(caption = "State Data: Change in Maternal Deaths per 1,000 Total Deaths, in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 (see Table 1).
-We used propensity weighting to account for differences between those states and the other 32 statesbased on the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008.
-We fit linear models with and without distances to abortion providers and an interaction term between those distances and restrictive legislation.
+We used propensity weighting to account for differences between those states and the other 32 states based on the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008 (the most recent presidential election in 2010).
+We fit linear models with and without distances to abortion providers and an interaction term between those distances and restrictive legislation as well as linear models with distance to abortion providers as the outcome to assess the relationship between restrictive legislation and acess to care.
 
@@
 We used abortion count data compiled by Wm.
-Robert Johnston and available at www.johnstonsarchive.net/policy/abortion.
+Robert Johnston, available at www.johnstonsarchive.net/policy/abortion.
 Abortion rates were calculated as number of abortions per female population.
@@
 
-For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, TRAP laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors.
-Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3/4 types of laws.
+For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute, which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, TRAP laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors.
+Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3 or 4 of these types of laws.
 
 We estimated distances to providers as the straight-line distances from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
+
+For the extension, data on state-level maternal, fetal and infant mortality rates as well as birth and fertility rates from the year 2010 were retrieved from the Centers for Disease Control and Prevention (CDC) WONDER Online Database.
 
@@
 Deaths Due to Remaining Pregnant" Demography (2021) 58(6):2019–2028.
+
+Centers for Disease Control and Prevention, National Center for Health Statistics.
+Compressed Mortality File 1999-2016 on CDC WONDER Online Database, released June 2017.
+Data are from the Compressed Mortality File 1999-2016 Series 20 No.
+2U, 2016, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program.
+Accessed at http://wonder.cdc.gov/cmf-icd10.html on Dec 15, 2021 9:43:02 PM
+
+United States Department of Health and Human Services (US DHHS), Centers of Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
+Linked Birth / Infant Death Records 2007-2018, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER On-line Database.
+Accessed at http://wonder.cdc.gov/lbd-current.html on Dec 15, 2021 8:57:30 PM
+
+United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics, Natality public-use data 2007-2020, on CDC WONDER Online Database, October 2021.
+Accessed at http://wonder.cdc.gov/natality-current.html on Dec 15, 2021 9:45:00 PM
+
+United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
+Fetal Deaths 2005-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER Online Database.
+Accessed at http://wonder.cdc.gov/fetal-deaths-current.html on Dec 15, 2021 9:45:49 PM

### 2021-12-16 report.Rmd
Update report.Rmd

@@
 ### Exposures
-In the replication, the exposures are a highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and distance to facilities performing abortions (in miles).
+In the replication, the exposures are a state-level highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and county-level distance to facilities performing abortions (in miles).
 For the extension, the exposures are the number of types of abortion legislation and the distance to facilities performing abortion (in miles).
@@
 For the extension, data on state-level maternal, fetal and infant mortality rates as well as birth and fertility rates from the year 2010 were retrieved from the Centers for Disease Control and Prevention (CDC) WONDER Online Database.
+State-level rates were used as opposed to county-level rates because more than 80% of these county-level rates were either suppressed or missing to maintain anonymity.
 
@@
 ### Replication using data from 18 states
+
 The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).
@@
 ### Expansion to data from all states
+
 When we repeated the analysis using all data available in all 50 states, we measured smaller but still significant reduction in the abortion rate, with associations of -0.216 and -0.361 without and with adjustment for distance to providers.
@@
 ## Sources
+
 Austin N, Harper S. "Assessing the impact of TRAP laws on abortion and women’s health in the USA: a systematic review" BMJ Sex Reprod Health, 2018; 44:128–134.
@@
 Accessed 12/14/21.
+
+Centers for Disease Control and Prevention, National Center for Health Statistics.
+Compressed Mortality File 1999-2016 on CDC WONDER Online Database, released June 2017.
+Data are from the Compressed Mortality File 1999-2016 Series 20 No.
+2U, 2016, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program.
+Accessed at http://wonder.cdc.gov/cmf-icd10.html on Dec 15, 2021 9:43:02 PM
 
@@
 
-Centers for Disease Control and Prevention, National Center for Health Statistics.
-Compressed Mortality File 1999-2016 on CDC WONDER Online Database, released June 2017.
-Data are from the Compressed Mortality File 1999-2016 Series 20 No.
-2U, 2016, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program.
-Accessed at http://wonder.cdc.gov/cmf-icd10.html on Dec 15, 2021 9:43:02 PM
+United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
+Fetal Deaths 2005-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER Online Database.
+Accessed at http://wonder.cdc.gov/fetal-deaths-current.html on Dec 15, 2021 9:45:49 PM
 
@@
 Accessed at http://wonder.cdc.gov/natality-current.html on Dec 15, 2021 9:45:00 PM
-
-United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
-Fetal Deaths 2005-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER Online Database.
-Accessed at http://wonder.cdc.gov/fetal-deaths-current.html on Dec 15, 2021 9:45:49 PM

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 table_no_dist <- tibble(Characteristic = "Count of Types of Restrictions",
-                  Beta = model_no_dist$coefficients[2],
+                  Beta = model_no_dist$coefficients[2]*1000,
                   "p-value" = c(p)) %>% 
@@
                                      "Distance/Restriction Interaction"),
-                  Beta = c(model_dist$coefficients[2], 
-                           model_dist$coefficients[3],
-                           model_dist$coefficients[4]),
+                  Beta = c(model_dist$coefficients[2]*1000, 
+                           model_dist$coefficients[3]*1000,
+                           model_dist$coefficients[4]*1000),
                   "p-value" = c(p_1, p_2, p_3)) %>%

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
+### Importance
+In the United States, many state legislatures have enacted a range of laws limiting access to abortions.
+These laws limit access to crucial healthcare and have impacts on mortality, fertility, and birth rates across the country.
+
+### Objectives
+To replicate the analysis of the association of abortion restrictions on abortion rates conducted by Brown et al.
+(2020), though not from a longitudinal perspective, to expand the analysis to a wider data set, and to extend the analysis to associations with maternal, fetal and infant mortality rates, as well as birth and fertility rates.
+
+### Design, Setting, and Particpants
+This study examined all publicly available county-level abortion rates in addition to state-level maternal, fetal and infant mortality rates as well as birth and fertility rates from the year 2010.
+The association between these rates, abortion legislation, and distance to nearest provider were evaluated using propensity score adjusted linear regression.
+
+### Exposures
+In the replication, the exposures are a state-level highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and county-level distance to facilities performing abortions (in miles).
+For the extension, the exposures are the number of types of abortion legislation and the distance to facilities performing abortion (in miles).
+
+### Main Outcomes and Measures
+In the replication and extension, the main outcomes and measures were county level abortion rates defined as abortions per 1,000 women per year.
+In the extension, the main outcomes and measures were state level fetal mortality rates defined as fetal deaths per 1,000 total deaths, infant mortality rates defined as infant deaths within the first year of life per 1,000 births, maternal mortality rates defined as deaths due to maternal diseases per 1,000 total deaths, birth rates defined as number of live births per 1,000 people, and fertility rates defined as number of live births per 1,000 women per year.
+
+### Results
+Our replication confirms the primary result of Brown et al.
+(2020).
+We found that highly restrictive legislation is associated with 0.814 fewer abortions per 1,000 women without adjusting for distance to providers, and 0.649 fewer abortions per 1,000 women when adjusting for distances to providers.
+These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, -0.92 to -0.04) and -0.44 (95% CI, -0.85 to -0.03), but our estimates were well within their reported 95% confidence intervals.
+
+### Conclusions and Relevance
+
+## Introduction
+
+In this paper we replicate and extend an analysis published in JAMA Network Open by Benjamin P.
+Brown, MD, MS; Luciana E.
+Hebert, PhD; Melissa Gilliam, MD, MPH;  and Robert Kaestner, PhD, entitled "Association of Highly Restrictive State Abortion Policies With Abortion Rates, 2000-2014."  Brown et al.
+use a binary classification of state legislative climates on abortion as either "highly restrictive" or not, and they assess the association with abortion rates using propensity weighted linear difference in differences models, based on longitudinal data from 18 states.
+
+We first reproduce the analysis of Brown et al.
+using data from the same 18 states.
+The data set from the Brown et al.
+article was not made public, and due to limitations of publicly available data, we have narrowed the scope of our replication to analyzing a single year.
+Specifically we analyzed data from 2010, primarily because this was the year within the original paper's time range for which we had the most detailed information on state legislation restricting abortions.
+Limitations of Brown et al.'s research included a lack of data on abortion rates in multiple states and a seemingly arbitrary classificaton of low vs.
+high legislative abortion restrictions.
+In order to improve the reliability and generalizability of their results, we chose to expand their analysis by using data from all 50 states, by using a more flexible integer variable for the extent of abortion restriction legislation in order, and by incorporating percentage of college graduates in the propensity score calculation.
+
+The original study addresses the immediate, intended consequences of the restrictive abortion legislation.
+In our paper, we extend the analysis to include associations of restrictive abortion legislation with fetal mortality, infant mortality, birth rate and fertility rate.
+Restricted access to abortion has been linked to adverse outcomes for mothers and children, including pre-term birth and low birth weight (Redd et al., 2021), infant mortality rates (Edwards, 2019), and maternal mortality (Stevenson, 2021).
+However, the effects of abortion legislation on these health outcomes are less well understood than the direct effect on number of abortions.
+The key findings of a meta-analysis on the effects of targeted restrictions on abortion providers (TRAP) were that "Certain TRAP laws may be associated with a decrease in state-level abortion rates" and "Additional research is required to better understand the impact of these laws on women’s health outcomes." (Austin and Harper, 2017)  We attempt to address this needd in the existing literature by adapting the methodology of Brown et al.
+to quantify the health impacts of restrictive abortion legislation.
+We also measure the association of these laws on birth rates, as restrictions on abortion have in some cases been promoted as an antedote to declining birth rates, recently in eastern Europe (Nikolova, 2020).
+Findings regarding whether restrictions on abortion actually increase birth rates, even if they decrease abortion rates, would therefore have important policy implications.
+
+## Study Design
+
+In our replication, we used propensity weighted linear models to assess the association of restrictive abortion legislation and abortion rates, using data from the same set of 18 states as Brown et al.
+(see Table 1).
+We used propensity weighting to account for differences between those states and the other 32 states based on the same set of county level covariates as Brown et al.: number of women, percentages of population by race, median household income, and percentage of voters who voted for the Democratic presidential candidate in 2008 (the most recent presidential election in 2010).
+We fit linear models with and without distances to abortion providers and an interaction term between those distances and restrictive legislation as well as linear models with distance to abortion providers as the outcome to assess the relationship between restrictive legislation and acess to care.
+
+## Participants, Variables and Data Sources
+
+We used abortion count data compiled by Wm.
+Robert Johnston, available at www.johnstonsarchive.net/policy/abortion.
+Abortion rates were calculated as number of abortions per female population.
+At least partial data was available for all 18 of the states included in the analysis by Brown et al., and for most of those states there was data for more than 80% of the counties (see Table 1).
+The percentages of counties reporting were similar to the percentages reporting in Brown et al.
+(see their Table 1).
+
+For assessment of state legislation restricting abortion, we used data from the Guttmacher Institute, which identified states with laws of the same four types used by Brown et al.: gestational age restrictions, TRAP laws, requirements for patients to make at least two trips to a clinic, and mandated parental involvement for patients who are minors.
+Following Brown et al.'s methodology, we classified states as "highly restrictive" if they had 3 or 4 of these types of laws.
+
+We estimated distances to providers as the straight-line distances from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
+
+For the extension, data on state-level maternal, fetal and infant mortality rates as well as birth and fertility rates from the year 2010 were retrieved from the Centers for Disease Control and Prevention (CDC) WONDER Online Database.
+State-level rates were used as opposed to county-level rates because more than 80% of these county-level rates were either suppressed or missing to maintain anonymity.
+
+## Results
+
+### Replication using data from 18 states
+
+The distribution of propensity scores for counties in states included in the analysis was similar to that for counties in states not included (see Figure 1).
+This suggests that the 18 states included in the analysis are not exceptionally different from those not included (in terms the demographic covariates listed above), and accordingly we expect the propensity weights to accurately adjust for the differences.
+
+Tables 3 and 4 show the main results from our replication.
+Without adjusting for distances to providers, we found that highly restrictive legislation was associated with a fewer abortions at a rate of -0.814 per 1,000 women without adjusting for distance to providers, and -0.649 when adjusting for distances to providers.
+These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, -0.92 to -0.04) and -0.44 (95% CI, -0.85 to -0.03), but our estimates were well within their reported 95% confidence intervals (see Brown et al.
+Table 3).
+This corroborates the main result of Brown et al.
+despite the use of completely different data sources.
+
+We did observe a decrease in abortion rates based on estimated distances to abortion providers.
+This decrease was statistically significant, but not practically significant at -3.4 x 10^-6 abortions per 1,000 women (see Table 4), compared to -0.02 in the analysis by Brown et al.
 
@@
 ## Expansion to Data from All States
+
+When we repeated the analysis using all data available in all 50 states, we measured smaller but still significant reduction in the abortion rate, with associations of -0.216 and -0.361 without and with adjustment for distance to providers.
+These again fall well within the 95% confidence interval of the estimates of Brown et al., further confirming their primary result, and suggesting that the propensity score weighting from the 18 states in their analysis to the entire country may be leading to an over-estimate of the effects.
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 \newpage
+## Limitations
+
+The original analysis and our replication are limited by the use of data from only 18 states.
+We address this shortcoming in our extension below.
+The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.
 
 ## Sources
+
+Austin N, Harper S. "Assessing the impact of TRAP laws on abortion and women’s health in the USA: a systematic review" BMJ Sex Reprod Health, 2018; 44:128–134.
+
+Edwards, E.
+"States pushing abortion bans have higher infant mortality rates".
+NBC News Website, 2019.
+Available at https://www.nbcnews.com/health/womens-health/states-pushing-abortion-bans-have-higher-infant-mortality-rates-n1008481.
+Accessed 12/14/21.
+
+Centers for Disease Control and Prevention, National Center for Health Statistics.
+Compressed Mortality File 1999-2016 on CDC WONDER Online Database, released June 2017.
+Data are from the Compressed Mortality File 1999-2016 Series 20 No.
+2U, 2016, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program.
+Accessed at http://wonder.cdc.gov/cmf-icd10.html on Dec 15, 2021 9:43:02 PM
+
+Nikolova, M.
+"When will conservatives learn that banning abortion does little to address demographic decline?" Emerging Europe website, 2020.
+Available at https://emerging-europe.com/after-hours/when-will-conservatives-learn-that-banning-abortion-does-little-to-address-demographic-decline/.
+Accessed 12/14/21.
+
+Redd, Sara K.; Stidham Hall, Kelli; Aswani, Monica S.; Sen, Bisakha; Wingate, Martha; Rice, Whitney S.
+"Variation in Restrictive Abortion Policies and Adverse Birth Outcomes in the United States from 2005 to 2015" Women's Health Issues, 2021 1-11.
+
+Stevenson, A. J. "The Pregnancy-Related Mortality Impact of a Total Abortion Ban in the United States: A Research Note on Increased
+Deaths Due to Remaining Pregnant" Demography (2021) 58(6):2019–2028.
+
+United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
+Fetal Deaths 2005-2019, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER Online Database.
+Accessed at http://wonder.cdc.gov/fetal-deaths-current.html on Dec 15, 2021 9:45:49 PM
+
+United States Department of Health and Human Services (US DHHS), Centers of Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics (DVS).
+Linked Birth / Infant Death Records 2007-2018, as compiled from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program, on CDC WONDER On-line Database.
+Accessed at http://wonder.cdc.gov/lbd-current.html on Dec 15, 2021 8:57:30 PM
+
+United States Department of Health and Human Services (US DHHS), Centers for Disease Control and Prevention (CDC), National Center for Health Statistics (NCHS), Division of Vital Statistics, Natality public-use data 2007-2020, on CDC WONDER Online Database, October 2021.
+Accessed at http://wonder.cdc.gov/natality-current.html on Dec 15, 2021 9:45:00 PM
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
+countydem$countyinanalysis <- ifelse(countydem$inanalysis == 1 & !is.na(countydem$AbortionRate2010), 1, 0) 
+countydem$countyinanalysislabel <- factor(countydem$countyinanalysis, levels = c(1, 0), labels = c("In Analysis", "Not in Analysis"))
+
+countydem$csinanalysis <- ifelse(sum(countydem$inanalysis + countydem$countyinanalysis) == 2, 1, 0)
+
 label(countydem$Women) = "Women"
@@
 
+statesummarydata <- countydem %>% filter(csinanalysis != 2) %>% group_by(state) %>% summarize(Women = mean(Women, na.rm = TRUE), White.pct = mean(White.pct, na.rm = TRUE), Black.pct = mean(Black.pct, na.rm = TRUE), Native.American.pct = mean(Native.American.pct, na.rm = TRUE), Asian.pct = mean(Asian.pct, na.rm = TRUE), Hispanic.pct = mean(Hispanic.pct, na.rm = TRUE), College.Grad.pct = mean(College.Grad.pct, na.rm = TRUE), Median.Income = mean(Median.Income, na.rm = TRUE), percentdem2008 = mean(percentdem2008, na.rm = TRUE), AvgDistancetoProvider = mean(distancetoprovider, na.rm = TRUE), inanalysis = mean(inanalysis, na.rm = TRUE))
+
+statesummarydata$inanalysislabel <- factor(statesummarydata$inanalysis, levels = c(1, 0), labels = c("In Analysis", "Not in Analysis"))
+
 ## Table 2 ###
 table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | inanalysislabel, 
-       data = countydem,
+       data = statesummarydata,
        overall = FALSE,

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 ##### Table 2 ####
@@
 countydem$csinanalysis <- ifelse(sum(countydem$inanalysis + countydem$countyinanalysis) == 2, 1, 0)
-
-label(countydem$Women) = "Women"
-label(countydem$White.pct) = "White"
-label(countydem$Black.pct) = "Black"
-label(countydem$Native.American.pct) = "Native American"
-label(countydem$Asian.pct) = "Asian"
-label(countydem$Hispanic.pct) = "Hispanic"
-label(countydem$Median.Income) = "Median Income"
-label(countydem$College.Grad.pct) = "College Graduates"
-label(countydem$percentdem2008) = "Vote Democratic"
-
-units(countydem$Women) = "No."
-units(countydem$White.pct) = "%"
-units(countydem$Black.pct) = "%"
-units(countydem$Native.American.pct) = "%"
-units(countydem$Asian.pct) = "%"
-units(countydem$Hispanic.pct) = "%"
-units(countydem$Median.Income) = "$"
-units(countydem$College.Grad.pct) = "%"
-units(countydem$percentdem2008) = "%"
 
@@
 statesummarydata$inanalysislabel <- factor(statesummarydata$inanalysis, levels = c(1, 0), labels = c("In Analysis", "Not in Analysis"))
+
+label(statesummarydata$Women) = "Women"
+label(statesummarydata$White.pct) = "White"
+label(statesummarydata$Black.pct) = "Black"
+label(statesummarydata$Native.American.pct) = "Native American"
+label(statesummarydata$Asian.pct) = "Asian"
+label(statesummarydata$Hispanic.pct) = "Hispanic"
+label(statesummarydata$Median.Income) = "Median Income"
+label(statesummarydata$College.Grad.pct) = "College Graduates"
+label(statesummarydata$percentdem2008) = "Vote Democratic"
+
+units(statesummarydata$Women) = "No."
+units(statesummarydata$White.pct) = "%"
+units(statesummarydata$Black.pct) = "%"
+units(statesummarydata$Native.American.pct) = "%"
+units(statesummarydata$Asian.pct) = "%"
+units(statesummarydata$Hispanic.pct) = "%"
+units(statesummarydata$Median.Income) = "$"
+units(statesummarydata$College.Grad.pct) = "%"
+units(statesummarydata$percentdem2008) = "%"
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 table_3 <- tibble(Characteristic = "Highly Restrictive Legislation",
-                  Beta = model_no_dist$coefficients[2]*1000,
-                  "p-value" = c('< 0.001')) %>% 
+                  Beta = summ_no_dist$coefficients[2,1]*1000,
+                  "95% CI lower" = summ_no_dist$coefficients[2,1]*1000-summ_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = summ_no_dist$coefficients[2,1]*1000+summ_no_dist$coefficients[2,2]*1000*1.96,
+                  "p-value" = '< 0.001') %>% 
   kbl(caption = "Change in Abortions per 1000 Women in Model without Distances", 
@@
                                      "Distance/Restriction Interaction"),
-                  Beta = c(model_dist$coefficients[2]*1000, model_dist$coefficients[3]*1000,
+                  Beta = c(model_dist$coefficients[2]*1000, 
+                           model_dist$coefficients[3]*1000,
                            model_dist$coefficients[4]*1000),
+                  "95% CI lower" = c(summ_dist$coefficients[2,1]*1000-summ_dist$coefficients[2,2]*1000*1.96,
+                                     summ_dist$coefficients[3,1]*1000-summ_dist$coefficients[3,2]*1000*1.96,
+                                     summ_dist$coefficients[4,1]*1000-summ_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(summ_dist$coefficients[2,1]*1000+summ_dist$coefficients[2,2]*1000*1.96,
+                                     summ_dist$coefficients[3,1]*1000+summ_dist$coefficients[3,2]*1000*1.96,
+                                     summ_dist$coefficients[4,1]*1000+summ_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 # Model Summaries (Note coefficients are per 1000 women)
+# Model Summaries (Note coefficients are per 1000 women)
 table_5 <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = all_state_model_no_dist$coefficients[2]*1000,
-                  "p-value" = c('< 0.001')) %>% 
+                  "95% CI lower" = all_state_model_no_dist$coefficients[2,1]*1000-all_state_model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = all_state_model_no_dist$coefficients[2,1]*1000+all_state_model_no_dist$coefficients[2,2]*1000*1.96,
+                  "p-value" = c('< 0.001')
+                  ) %>% 
   kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
@@
                            all_state_model_dist$coefficients[4]*1000),
+                  "95% CI lower" = c(all_state_model_dist$coefficients[2,1]*1000-all_state_model_dist$coefficients[2,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[3,1]*1000-all_state_model_dist$coefficients[3,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[4,1]*1000-all_state_model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(all_state_model_dist$coefficients[2,1]*1000+all_state_model_dist$coefficients[2,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[3,1]*1000+all_state_model_dist$coefficients[3,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[4,1]*1000+all_state_model_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

--- 2021-12-16 paper_replication.Rmd
@@
 # Model Summaries (Note coefficients are per 1000 women)
+# Model Summaries (Note coefficients are per 1000 women)
 table_5 <- tibble(Characteristic = "Count of Types of Restrictions",
                   Beta = all_state_model_no_dist$coefficients[2]*1000,
-                  "p-value" = c('< 0.001')) %>% 
-  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model without Distances", 
+                  "95% CI lower" = all_state_model_no_dist$coefficients[2,1]*1000-all_state_model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = all_state_model_no_dist$coefficients[2,1]*1000+all_state_model_no_dist$coefficients[2,2]*1000*1.96,
+                  "p-value" = c('< 0.001')
+                  ) %>% 
+  kbl(caption = "Counties from All States: Change in Abortions per 1000 Women in Model without Distances", 
           digits = 3) %>%
@@
                            all_state_model_dist$coefficients[4]*1000),
+                  "95% CI lower" = c(all_state_model_dist$coefficients[2,1]*1000-all_state_model_dist$coefficients[2,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[3,1]*1000-all_state_model_dist$coefficients[3,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[4,1]*1000-all_state_model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(all_state_model_dist$coefficients[2,1]*1000+all_state_model_dist$coefficients[2,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[3,1]*1000+all_state_model_dist$coefficients[3,2]*1000*1.96,
+                                     all_state_model_dist$coefficients[4,1]*1000+all_state_model_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c('< 0.001', '< 0.001', p)) %>%
-  kbl(caption = "All States Data: Change in Abortions per 1000 Women in Model with Distances") %>%
+  kbl(caption = "Counties from All States: Change in Abortions per 1000 Women in Model with Distances") %>%
   kable_styling(latex_options = 'hold_position')
@@
                   Beta = model_no_dist$coefficients[2]*1000,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
                   "p-value" = c(p)) %>% 
@@
                            model_dist$coefficients[4]*1000),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c(p_1, p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
                   "p-value" = c(p)) %>% 
@@
                            model_dist$coefficients[4]),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c(p_1, p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
                   "p-value" = '< 0.001') %>% 
@@
                            model_dist$coefficients[4]),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1000*1.96,
+                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1000*1.96,
+                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1000*1.96),
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
                   "p-value" = p) %>% 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
-## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ```
-## Expansion to Data from All States
+### Expansion to Data from All States
 
@@
 ```
-## Extension to Infant Death Rate, Birth Rate, and Fertility Rate
+### Extension to Infant Death Rate, Birth Rate, and Fertility Rate
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = c(p)) %>% 
@@
                            model_dist$coefficients[4]),
-                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1000*1.96,
-                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1000*1.96,
-                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1000*1.96),
-                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1000*1.96,
-                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1000*1.96,
-                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1.96),
                   "p-value" = c(p_1, p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = '< 0.001') %>% 
@@
                            model_dist$coefficients[4]),
-                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1000*1.96,
-                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1000*1.96,
-                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1000*1.96),
-                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1000*1.96,
-                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1000*1.96,
-                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1000*1.96),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1.96),
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1000*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1000*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = p) %>% 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
-  #plot of distribution of propensity scores 
-  # p1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
-  # xlab('Probability of Being in the Analysis') +
-  # ggtitle('Propensity Score Distribution by Group') +
-  # scale_fill_discrete('') +
-  # theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  # ylab("Density")
-  # print(p1)
-
   #run model
@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = c(p)) %>% 
@@
                            model_dist$coefficients[4]),
-                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1.96,
-                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1.96,
-                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1.96),
-                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1.96,
-                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1.96,
-                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1.96),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]-model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]-model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]-model_dist$coefficients[4,2]*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]+model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]+model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]+model_dist$coefficients[4,2]*1.96),
                   "p-value" = c(p_1, p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = '< 0.001') %>% 
@@
                            model_dist$coefficients[4]),
-                  "95% CI lower" = c(model_dist$coefficients[2,1]*1000-model_dist$coefficients[2,2]*1.96,
-                                     model_dist$coefficients[3,1]*1000-model_dist$coefficients[3,2]*1.96,
-                                     model_dist$coefficients[4,1]*1000-model_dist$coefficients[4,2]*1.96),
-                  "95% CI upper" = c(model_dist$coefficients[2,1]*1000+model_dist$coefficients[2,2]*1.96,
-                                     model_dist$coefficients[3,1]*1000+model_dist$coefficients[3,2]*1.96,
-                                     model_dist$coefficients[4,1]*1000+model_dist$coefficients[4,2]*1.96),
+                  "95% CI lower" = c(model_dist$coefficients[2,1]-model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]-model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]-model_dist$coefficients[4,2]*1.96),
+                  "95% CI upper" = c(model_dist$coefficients[2,1]+model_dist$coefficients[2,2]*1.96,
+                                     model_dist$coefficients[3,1]+model_dist$coefficients[3,2]*1.96,
+                                     model_dist$coefficients[4,1]+model_dist$coefficients[4,2]*1.96),
                   "p-value" = c('< 0.001', p_2, p_3)) %>%
@@
                   Beta = model_no_dist$coefficients[2],
-                  "95% CI lower" = model_no_dist$coefficients[2,1]*1000-model_no_dist$coefficients[2,2]*1.96,
-                  "95% CI upper" = model_no_dist$coefficients[2,1]*1000+model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI lower" = model_no_dist$coefficients[2,1]-model_no_dist$coefficients[2,2]*1.96,
+                  "95% CI upper" = model_no_dist$coefficients[2,1]+model_no_dist$coefficients[2,2]*1.96,
                   "p-value" = p) %>% 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
-When we repeated the analysis using all data available in all 50 states, we measured smaller but still significant reduction in the abortion rate, with associations of -0.216 and -0.361 without and with adjustment for distance to providers.
-These again fall well within the 95% confidence interval of the estimates of Brown et al., further confirming their primary result, and suggesting that the propensity score weighting from the 18 states in their analysis to the entire country may be leading to an over-estimate of the effects.
+When we repeated the analysis using all data available in all 50 states, we measured a significant reduction in the abortion rate, with associations per type of restriction in place of -0.216 and -0.361 without and with adjustment for distance to providers (see Tables 5 and 6).
+These results align with the results of Brown et al.
+and the results of our replication, and suggest that the propensity score weighting from the 18 states in their analysis to the entire country is accurately accounting for differences between those 18 states and the others.
+We repeated the analysis of abortion rate using state level data from all states(see Tables 7 and 8), and found stronger but comparable associations, further confirming the primary resutls of Brown et al.
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
   units(countydem$percentdem2008) = "%"
-
-  # p_value_t_test <- function(x, ...){
-  #   y <- unlist(x[1:2])
-  #   g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
-  #   p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
-  #   p_value <- ifelse(p_value == '0', '< 0.001', p_value)
-  # }
-  # 
-  # in_analysis_comp_table <- table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasdatalabel, 
-  #      data = countydem,
-  #      extra.col = list(`P-value` = p_value_t_test),
-  #      footnote = 'P-values are derived by two sample t-tests.'
-  #      )
 
@@
 
-  # print(table1(~ AvgWomen + AvgWhite.pct + AvgBlack.pct + AvgNative.American.pct + AvgAsian.pct + AvgHispanic.pct + AvgMedian.Income + AvgCollege.Grad.pct + AvgDem2008.pct | hasdatalabel, 
-  #      data = statedem,
-  #      extra.col = list(`P-value` = p_value_t_test),
-  #      footnote = 'P-values are derived by two sample t-tests.'
-  #      ))
-
   #calculate propensity scores
@@
   df$ps <- p.score
-
-  #plot of distribution of propensity scores
-
-  # plot1 <- ggplot(df, aes(x = ps, group = hasdatalabel)) + geom_density(aes(fill = hasdatalabel), alpha = 0.2) +
-  #   xlab('Probability of Being in the Analysis') +
-  #   ggtitle('Propensity Score Distribution by Group') +
-  #   scale_fill_discrete('') +
-  #   theme(legend.position = 'bottom', legend.direction = 'vertical') +
-  #   ylab("Density")
-  # print(plot1)
 
@@
 ### Extension to Infant Death Rate, Birth Rate, and Fertility Rate
+We found positive associations between abortion restrictions and both birth rates and fertility rates (see Tables 9-12), however these associations were not statistically significant after adjusting for distance to providers, with p-values > 0.5.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 and the results of our replication, and suggest that the propensity score weighting from the 18 states in their analysis to the entire country is accurately accounting for differences between those 18 states and the others.
-We repeated the analysis of abortion rate using state level data from all states(see Tables 7 and 8), and found stronger but comparable associations, further confirming the primary resutls of Brown et al.
+We repeated the analysis of abortion rate using state level data from all states(see Tables 7 and 8), and found stronger but comparable associations, further confirming the primary results of Brown et al.
 
@@
 \newpage
+
 ## Limitations
 
-The original analysis and our replication are limited by the use of data from only 18 states.
-We address this shortcoming in our extension below.
+The original analysis and our replication of this analysis are limited by the use of data from only 18 states.
+We addressed this shortcoming in our extension, though our extension still does not have data for all states and/or all counties.
+Further, our replication of the original analysis as well as our extensions and expansions of said analysis are limited by the fact that state-level legislative data is only available for the years 2000, 2010, and 2020.
+More detailed legislative information is available for more recent years (roughly 2018 to the present), but abortion count data for these years is not yet released.
+Further analyses with more recent, detailed legislative data and abortion counts from more recent years would greatly add to the present analyses and could also allow for a longitudinal approach.
 The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 These effects are somewhat larger in magnitude than the effects calculated by Brown et al., which were -0.48 (95% CI, -0.92 to -0.04) and -0.44 (95% CI, -0.85 to -0.03), but our estimates were well within their reported 95% confidence intervals.
+Infant mortality rates were positively associated with abortion restrictions, and these associations were statistically significant even after adjusting for distance from providers.
+The maternal and fetal mortality rates were not statistically significant.
 
@@
 We found positive associations between abortion restrictions and both birth rates and fertility rates (see Tables 9-12), however these associations were not statistically significant after adjusting for distance to providers, with p-values > 0.5.
+Infant mortality rates were positively associated with abortion restrictions, and these associations were statistically significant even after adjusting for distance from providers (see Tables 13 and 14) .
+The maternal and fetal mortality rates were not statistically significant (see Tables 15-18).
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
-The original analysis and our replication of this analysis are limited by the use of data from only 18 states.
+The main limitations of these analyses are due to missing data.
+Both the original analysis and our replication of this analysis are limited by the use of data from only 18 states.
 We addressed this shortcoming in our extension, though our extension still does not have data for all states and/or all counties.
@@
 More detailed legislative information is available for more recent years (roughly 2018 to the present), but abortion count data for these years is not yet released.
-Further analyses with more recent, detailed legislative data and abortion counts from more recent years would greatly add to the present analyses and could also allow for a longitudinal approach.
-The original analysis and our own are limited by the calculating abortion rates using total number of women, rather than number of women of reproductive age.
+Further analyses with more recent, detailed legislative data and abortion counts from more recent years would greatly add to the present analyses and could also allow for a more longitudinal approach.
+The original analysis and our own are also limited by the fact that abortion rates are calculated using total number of women, rather than number of women of reproductive age - this is due to the lack of data on population statistics stratified by age.
+Further, the data used in our extension is only by state, rather than by county, as previously discussed, due to suppression of data for the purpose of preserving anonymity.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 ---
+## Abstract
 ### Importance
@@
 ### Conclusions and Relevance
+Our analysis confirms the results of Brown et al.
+showing a negative association between restrictive abortion legislation and abortion rates based on both the 18 states they analyized and based on data from all 50 states.
+We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 showing a negative association between restrictive abortion legislation and abortion rates based on both the 18 states they analyized and based on data from all 50 states.
-We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates.
+We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates, and showed that restrictions do not have a significant association with birth rates.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

--- 2021-12-16 paper_replication.Rmd
@@
 ---
-title: "paper_replication"
+title: "Replication and Extension of Analyses Regarding Effects of Highly Restrictive Abortion Policies"
 output: pdf_document
@@
 ### Exposures
-In the replication, the exposures are a state-level highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and county-level distance to facilities performing abortions (in miles).
-For the extension, the exposures are the number of types of abortion legislation and the distance to facilities performing abortion (in miles).
+In the replication, the exposures are state-level highly restrictive abortion legislation as defined by 3 or more out of 4 common types of abortion restrictions and county-level distance to facilities performing abortions (in miles).
+For the extension, the exposures are the number of types of abortion legislation and the state-level average distance to facilities performing abortion (in miles).
 
@@
 Infant mortality rates were positively associated with abortion restrictions, and these associations were statistically significant even after adjusting for distance from providers.
-The maternal and fetal mortality rates were not statistically significant.
+The maternal and fetal mortality rates as well as the birth rates were not statistically significant after adjusting for distance from providers.
 
@@
 showing a negative association between restrictive abortion legislation and abortion rates based on both the 18 states they analyized and based on data from all 50 states.
-We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates.
+We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates, while restrictions do not have a significant association with birth rates or maternal or fetal mortality rates after adjusting for distance to providers.
 
@@
 We estimated distances to providers as the straight-line distances from the population centroids of counties of residence (courtesy of the United States Census Bureau) to the location of the nearest abortion clinic (courtesy of Safe Place Project).
+
+Data on the percentage of democratic votes by county in the most recent (2008) presidential election was retrieved from the Harvard Dataverse.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 
+### Conclusions & Relevance
+Our analysis confirms the results of Brown et al.
+showing a negative association between restrictive abortion legislation and abortion rates based on both the 18 states they analyized and based on data from all 50 states.
+We furthered the analysis to additional outcomes and found positive associations between restrictions and infant death rates, while restrictions do not have a significant association with birth rates or maternal or fetal mortality rates after adjusting for distance to providers.
+Ultimately, this suggests that while highly restrictive legislation might significantly decrease abortion rates, as shown by Brown et al., such legislation does not actually significantly increase birth rates and actually increases infant death rates.
+This conclusion ultimately suggests that policies enforcing strict legislation on abortion restrictions would most likely be ineffective at increasing populations - while abortions decrease, births do not increase, suggesting that pregnancies may be terminated or prevented in other ways, and infant deaths actually increase.
+All together, highly restrictive legislation appears to have negative consequences on populations across the United States: such legislation reduces accessibility to important healthcare, may cause less-safe abortions that are not performed in medical clinics, and may result in more babies being born who are very sick.
+
 ## Limitations
@@
 Further, the data used in our extension is only by state, rather than by county, as previously discussed, due to suppression of data for the purpose of preserving anonymity.
+As a final note, the nature of abortion, mortality, fertility, and birth data is limited by the fact that not all of these occurrences are reported, and, if they are, not all ofthem are reported with complete accuracy (for example, it's possible that the abortion rates is much higher than those reported in the datasets and that women do not report all abortions) due to the sensitive nature of such data.
 

### 2021-12-16 paper_replication.Rmd
Update paper_replication.Rmd

@@
 We did observe a decrease in abortion rates based on estimated distances to abortion providers.
-This decrease was statistically significant, but not practically significant at -3.4 x 10^-6 abortions per 1,000 women (see Table 4), compared to -0.02 in the analysis by Brown et al.
+This decrease was statistically significant significant at -0.00554 abortions per 1,000 women (see Table 4), compared to -0.02 in the analysis by Brown et al.
 
